### 2021-11-30 Fish Data Analysis/Ji_Zhou_Data.Rmd

---
title: "Fish Data For Ji Zhou"
output: html_notebook
---

```{r}
library(tidyverse)
library(ggplot2)
library(car)
library(viridis)
library(boot)
```

```{r}
comp_data <- read.csv("Fish_Comp_Values.csv")
comp_data <- na.omit(comp_data)

comp_data <- comp_data %>% mutate(Flow = ifelse(Flow == "0", "Flow 0", "Flow 2")) %>%
                           mutate(Ablation = ifelse(Ablation == "N", "No Ablation", "Ablated")) %>%
                           mutate(Darkness = ifelse(Darkness == "N", "Light", "Dark")) %>%
                           mutate(Heading_Diff = abs(Heading_Diff)) %>%
                           filter(Distance <= 4) %>%
                           filter(Speed_Diff <= 6) %>% 
                           mutate(quarter_heading_diff = abs(abs(Heading_Diff-90)-90)) %>%
                           mutate(Is_Aligned = ifelse(Heading_Diff < 30, 1, 0)) %>%
                           mutate(Is_Reversed = ifelse(Heading_Diff > 150, 1, 0)) %>%
                           mutate(Angle = fold_angle_0_360_to_0_180(Angle))

indiv_data <- indiv_data %>% mutate(Flow = ifelse(Flow == "0", "Flow 0", "Flow 2")) %>%
                             mutate(Ablation = ifelse(Ablation == "N", "No Ablation", "Ablated")) %>%
                             mutate(Darkness = ifelse(Darkness == "N", "Light", "Dark"))

```

Tank Size: 60cm long and 25cm wide

Fish Size:

  Small Fish: 
    Length: 
    Width:
    Height:

  Big Fish: 
    Length: 
    Width:
    Height:

Flow: 10.16 cm/s

```{r}
comp_data %>% group_by(Flow, Ablation, Darkness) %>% summarise(mean_Dist = round(mean(Distance),2), 
                                                                sd_Dist = round(sd(Distance),2))
```

NND:
  Still Mean: 1.87
  Still SD:   0.89

  Flow Mean:  1.90
  Flow SD:    1.00

```{r}
indiv_data %>% group_by(Flow, Ablation, Darkness) %>% summarise(mean_Dist = round(mean(Speed),2), 
                                                                sd_Dist = round(sd(Speed),2))
```

Speed:
  Still Mean: 1.42
  Still SD:   0.83

  Flow Mean:  2.38
  Flow SD:    0.66

```{r}
indiv_data %>% group_by(Flow, Ablation, Darkness) %>% summarise(mean_Dist = round(mean(TB_Frequency),2), 
                                                                sd_Dist = round(sd(TB_Frequency),2))
```

Tailbeat Frequency:
  Still Mean: 2.29
  Still SD:   1.15

  Flow Mean:  2.82
  Flow SD:    1.43