### 2021-11-30 Fish Data Analysis/comp_fish_simple_models.Rmd

---
title: "Fish Comparisons Simple Models"
output: html_notebook
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(car)
library(viridis)
library(boot)
```

So here we're going to do some simpler model selection things with the data using GLMs instead of GAMs

Sets up the functions that will be used later 
```{r}
rad2deg <- function(rad) {(rad * 180) / (pi)}
deg2rad <- function(deg) {(deg * pi) / (180)}
round_any <- function(x, accuracy, f=round){f(x/ accuracy) * accuracy}
ang_mean <- function(x){rad2deg(atan(mean(sin(deg2rad(x)))/mean(cos(deg2rad(x)))))}

fold_angle_0_360_to_0_180 <- function(x){abs(abs(x-180)-180)}
```

Reads in the data and alters it as needed 
```{r}
comp_data <- read.csv("Fish_Comp_Values.csv")
comp_data <- na.omit(comp_data)

comp_data <- comp_data %>% mutate(Flow = ifelse(Flow == "0", "Flow 0", "Flow 2")) %>%
                           mutate(Ablation = ifelse(Ablation == "N", "No Ablation", "Ablated")) %>%
                           mutate(Darkness = ifelse(Darkness == "N", "Light", "Dark")) %>%
                           mutate(Heading_Diff = abs(Heading_Diff)) %>%
                           filter(Distance <= 4) %>%
                           filter(Speed_Diff <= 6) %>% 
                           mutate(quarter_heading_diff = abs(abs(Heading_Diff-90)-90)) %>%
                           mutate(Is_Aligned = ifelse(Heading_Diff < 30, 1, 0)) %>%
                           mutate(Is_Reversed = ifelse(Heading_Diff > 150, 1, 0)) %>%
                           mutate(Angle = fold_angle_0_360_to_0_180(Angle))

sum_comp_data <- comp_data %>% mutate(X_Distance = round_any(X_Distance,0.25), 
                                      Y_Distance = round_any(abs(Y_Distance),0.25)) %>%
                               group_by(Flow,Ablation,Darkness,X_Distance,Y_Distance) %>%
                               summarise(Speed_Diff = mean(Speed_Diff),
                                         Heading_Diff = ang_mean(Heading_Diff),
                                         Sync = mean(Sync),
                                         Quarter_Heading_Diff = mean(quarter_heading_diff),
                                         Is_Aligned = mean(Is_Aligned),
                                         Is_Reversed = mean(Is_Reversed))

```

#Basic Stats

```{r}
speed_anova <- glm(Speed_Diff ~ Flow + Ablation + Darkness + Flow:Ablation + Flow:Darkness, data = comp_data)

Anova(speed_anova)

heading_anova <- glm(quarter_heading_diff ~ Flow + Ablation + Darkness + Flow:Ablation + Flow:Darkness, data = comp_data)

Anova(heading_anova)

sync_anova <- glm(Sync ~ Flow + Ablation + Darkness + Flow:Ablation + Flow:Darkness, data = comp_data)

Anova(sync_anova)

dist_anova <- glm(Distance ~ Flow + Ablation + Darkness + Flow:Ablation + Flow:Darkness, data = comp_data)

Anova(dist_anova)
```

```{r}
ggplot(comp_data, aes(x = Flow, y = Speed_Diff, fill = paste(Ablation,Darkness,sep="-")))+
  geom_boxplot(outlier.shape = NA) +
  ylim(0,3) +
  theme_light()

ggplot(comp_data, aes(x = Flow, y = quarter_heading_diff, fill = paste(Ablation,Darkness,sep="-")))+
  geom_boxplot(outlier.shape = NA) +
  ylim(0,90) +
  theme_light()

ggplot(comp_data, aes(x = Flow, y = Sync, fill = paste(Ablation,Darkness,sep="-")))+
  geom_boxplot(outlier.shape = NA) +
  ylim(0,1) +
  theme_light()

ggplot(comp_data, aes(x = Flow, y = Distance, fill = paste(Ablation,Darkness,sep="-")))+
  geom_boxplot(outlier.shape = NA) +
  ylim(0,4) +
  theme_light()
```

Now we add in distance and angle

```{r}
speed_anova_da <- glm(Speed_Diff ~ Distance*Angle*(Flow + Ablation + Darkness + Flow:Ablation + Flow:Darkness), data = comp_data)

Anova(speed_anova_da)

heading_anova_da <- glm(quarter_heading_diff ~ Distance*Angle*(Flow + Ablation + Darkness + Flow:Ablation + Flow:Darkness), data = comp_data)

Anova(heading_anova_da)

sync_anova_da <- glm(Sync ~ Distance*Angle*(Flow + Ablation + Darkness + Flow:Ablation + Flow:Darkness), data = comp_data)

Anova(sync_anova_da)
```

```{r}
ggplot(comp_data, aes(x = Distance, y = Speed_Diff, color = paste(Ablation,Darkness,sep="-"), fill = paste(Ablation,Darkness,sep="-")))+
  geom_smooth(method = "glm", formula = y ~ poly(x,4))+
  facet_wrap(~ Flow) +
  theme_light()

ggplot(comp_data, aes(x = Distance, y = quarter_heading_diff, color = paste(Ablation,Darkness,sep="-"), fill = paste(Ablation,Darkness,sep="-")))+
  geom_smooth(method = "glm", formula = y ~ poly(x,4))+
  facet_wrap(~ Flow) +
  theme_light()

ggplot(comp_data, aes(x = Distance, y = Sync, color = paste(Ablation,Darkness,sep="-"), fill = paste(Ablation,Darkness,sep="-")))+
  geom_smooth(method = "glm", formula = y ~ poly(x,4))+
  facet_wrap(~ Flow) +
  theme_light()
```
```{r}
ggplot(comp_data, aes(x = Angle, y = Speed_Diff, color = paste(Ablation,Darkness,sep="-"), fill = paste(Ablation,Darkness,sep="-")))+
  geom_smooth(method = "glm", formula = y ~ poly(x,4))+
  facet_wrap(~ Flow) +
  theme_light()

ggplot(comp_data, aes(x = Angle, y = quarter_heading_diff, color = paste(Ablation,Darkness,sep="-"), fill = paste(Ablation,Darkness,sep="-")))+
  geom_smooth(method = "glm", formula = y ~ poly(x,4))+
  facet_wrap(~ Flow) +
  theme_light()

ggplot(comp_data, aes(x = Angle, y = Sync, color = paste(Ablation,Darkness,sep="-"), fill = paste(Ablation,Darkness,sep="-")))+
  geom_smooth(method = "glm", formula = y ~ poly(x,4))+
  facet_wrap(~ Flow) +
  theme_light()
```

This creates a dataframe that is used for making model predictions and graphing them
```{r}
d <- seq(from = 0, to = 4, by = 0.1)
a <- seq(from = 0, to = 180, by = 10)

flows <- c("Flow 0", "Flow 2")
ablation <- c("No Ablation", "Ablated")
dark <- c("Light","Dark")

predict_df_da <- expand.grid(Distance = d, Angle = a, Flow = flows, Ablation = ablation, Darkness = dark)
predict_df_da <- predict_df_da %>% mutate(X_Distance = Distance*(cos(deg2rad(Angle))), Y_Distance = Distance*(sin(deg2rad(Angle))))

predict_df_da <- predict_df_da %>% filter(!(Ablation == "Ablated" & Darkness == 'Dark'))
predict_df_da <- na.omit(predict_df_da)
```

### Speed Models

So I want to make a model that tries to include both distance and angle. What I really need is to figure out what power works best for angle and distance, and then include that. The issue is stepwise selection can just remove one of them entirely. So I'm going to use cross validation to select the power for both of them, and then use that. I will also calculate the AICs to help in this decision making process

```{r, warnings=F}
set.seed(7)

comp_data_speed_model <- comp_data %>% select(c(Flow,Darkness,Ablation,Angle,Distance,Speed_Diff))

max_poly <- 10

speed_dist_cv_error_10 <- rep(0,max_poly)
speed_dist_AIC_10 <- rep(0,max_poly)

for (i in 1:max_poly){
  speed_dist_fit <- glm(Speed_Diff ~ poly(Distance,i)*(Ablation+Flow+Darkness+Ablation:Flow+Darkness:Flow),
                        data = comp_data_speed_model)
  speed_dist_cv_error_10[i] <- cv.glm(comp_data_speed_model, speed_dist_fit, K = 10)$delta[1]
  speed_dist_AIC_10[i] <- AIC(speed_dist_fit)
}

speed_angle_cv_error_10 <- rep(0,max_poly)
speed_angle_AIC_10 <- rep(0,max_poly)

for (i in 1:max_poly){
  speed_angle_fit <- glm(Speed_Diff ~ poly(Angle,i)*(Ablation+Flow+Darkness+Ablation:Flow+Darkness:Flow),
                         data = comp_data_speed_model)
  speed_angle_cv_error_10[i] <- cv.glm(comp_data_speed_model, speed_angle_fit, K = 10)$delta[1]
  speed_angle_AIC_10[i] <- AIC(speed_angle_fit)
}

speed_poly_plot_df <- data.frame(c(seq(max_poly),seq(max_poly)),
                                 c(speed_dist_cv_error_10,speed_angle_cv_error_10),
                                 c(speed_dist_AIC_10,speed_angle_AIC_10),
                                 c(rep("Distance",max_poly),rep("Angle",max_poly)))

colnames(speed_poly_plot_df) <- c("Degree","Error","AIC","Predictor")

speed_poly_plot_df <- speed_poly_plot_df %>% group_by(Predictor) %>%
                                             mutate(minError = min(Error),minAIC = min(AIC)) %>%
                                             ungroup() %>%
                                             mutate(isMinEror = ifelse(Error == minError,3,1),isMinAIC = ifelse(AIC == minAIC,3,1))

ggplot(speed_poly_plot_df, aes(x = Degree, y = Error, color = Predictor))+
  geom_point(size = speed_poly_plot_df$isMinEror)+
  geom_line()+
  theme_light()+
  facet_wrap(~ Predictor, scales = "free") +
  scale_size(guide = "none")

ggplot(speed_poly_plot_df, aes(x = Degree, y = AIC, color = Predictor))+
  geom_point(size = speed_poly_plot_df$isMinAIC)+
  geom_line()+
  theme_light()+
  facet_wrap(~ Predictor, scales = "free") +
  scale_size(guide = "none")
```
Now let's use those polynomials to create the actual model

While they not be the lowest, the big improvements happen at 4 for Distance, and 2 for Angle

```{r}
speed_diff_glm <- glm(Speed_Diff ~ poly(Distance,4)*poly(Angle,2)*(Ablation+Flow+Darkness+Ablation:Flow+Darkness:Flow),
                        data = comp_data_speed_model)

summary(speed_diff_glm)
```

Now let's make some predictions

```{r}
speed_pred <- predict_df_da %>% mutate(Speed_Diff = predict(speed_diff_glm,predict_df_da))

comp_data <- comp_data %>% mutate(Round_Dist = as.factor(round_any(Distance,1)), Round_Angle = as.factor(round_any(Angle,30)))

ggplot()+
  geom_boxplot(data = comp_data, aes(x = Round_Dist, y = Speed_Diff))+
  facet_wrap(~ Flow + Ablation + Darkness) +
  theme_light()

round_dist_aov <- aov(Speed_Diff ~ Round_Dist*(Flow + Ablation + Darkness + Flow:Ablation + Flow:Darkness), data = comp_data)
Anova(round_dist_aov)

ggplot()+
  geom_boxplot(data = comp_data, aes(x = Round_Angle, y = Speed_Diff))+
  facet_wrap(~ Flow + Ablation + Darkness) +
  theme_light()

round_angle_aov <- aov(Speed_Diff ~ Round_Angle*(Flow + Ablation + Darkness + Flow:Ablation + Flow:Darkness), data = comp_data)
Anova(round_angle_aov)

# ggplot()+
#   geom_point(data = comp_data %>% filter(Speed_Diff <= 2), aes(x = Distance, y = Speed_Diff), alpha = 0.1)+
#   geom_smooth(data = speed_pred, aes(x = Distance, y = Speed_Diff))+
#   facet_wrap(~ Flow + Ablation + Darkness) +
#   theme_light()

# ggplot()+
#   geom_density_2d_filled(data = comp_data %>% filter(Speed_Diff <= 2), aes(x = Distance, y = Speed_Diff), contour_var = "ndensity")+
#   geom_smooth(data = speed_pred, aes(x = Distance, y = Speed_Diff, color = "red"))+
#   facet_wrap(~ Flow + Ablation + Darkness) +
#   theme_light()

# ggplot()+
#   geom_point(data = comp_data %>% filter(Speed_Diff <= 2), aes(x = Angle, y = Speed_Diff), alpha = 0.1)+
#   geom_smooth(data = speed_pred, aes(x = Angle, y = Speed_Diff))+
#   facet_wrap(~ Flow + Ablation + Darkness) +
#   theme_light()

# ggplot()+
#   geom_density_2d_filled(data = comp_data %>% filter(Speed_Diff <= 2), aes(x = Angle, y = Speed_Diff), contour_var = "ndensity")+
#   geom_smooth(data = speed_pred, aes(x = Angle, y = Speed_Diff, color = "red"))+
#   facet_wrap(~ Flow + Ablation + Darkness) +
#   theme_light()

ggplot(data = comp_data, aes(x = Angle, y = Distance, z = Speed_Diff))+
  stat_summary_2d() +
  facet_wrap(~ Flow + Ablation + Darkness) +
  scale_fill_viridis(direction = -1) +
  theme_light()
```
You know let's just try stepwise as well. There's jsut way too many thing in that possible model

```{r}

speed_m_all <- glm(Speed_Diff ~ Distance*Angle*(Ablation+Flow+Darkness+Ablation:Flow+Darkness:Flow) +
                                I(Distance^2)*I(Angle^2)*(Ablation+Flow+Darkness+Ablation:Flow+Darkness:Flow)+
                                I(Distance^3)*I(Angle^3)*(Ablation+Flow+Darkness+Ablation:Flow+Darkness:Flow)+
                                I(Distance^4)*I(Angle^4)*(Ablation+Flow+Darkness+Ablation:Flow+Darkness:Flow),
                        data = comp_data_speed_model)

speed_m_none <- glm(Speed_Diff ~ 1, data = comp_data_speed_model)

speed_m_both <- step(speed_m_none, direction = "both", scope = formula(speed_m_all), trace = F)

summary(speed_m_both)
```

### Sync Models

Now let's try the same thing for sync values (Rayleigh's R)

```{r, warnings=F}
set.seed(7)

comp_data_sync_model <- comp_data %>% select(c(Flow,Darkness,Ablation,Angle,Distance,Sync))

max_poly <- 10

sync_dist_cv_error_10 <- rep(0,max_poly)
sybc_dist_AIC_10 <- rep(0,max_poly)

for (i in 1:max_poly){
  sync_dist_fit <- glm(Sync ~ poly(Distance,i)*(Ablation+Flow+Darkness+Ablation:Flow+Darkness:Flow),
                        data = comp_data_sync_model)
  sync_dist_cv_error_10[i] <- cv.glm(comp_data_sync_model, sync_dist_fit, K = 10)$delta[1]
  sybc_dist_AIC_10[i] <- AIC(sync_dist_fit)
}

sync_angle_cv_error_10 <- rep(0,max_poly)
sync_angle_AIC_10 <- rep(0,max_poly)

for (i in 1:max_poly){
  sync_angle_fit <- glm(Sync ~ poly(Angle,i)*(Ablation+Flow+Darkness+Ablation:Flow+Darkness:Flow),
                         data = comp_data_sync_model)
  sync_angle_cv_error_10[i] <- cv.glm(comp_data_sync_model, sync_angle_fit, K = 10)$delta[1]
  sync_angle_AIC_10[i] <- AIC(sync_angle_fit)
}

sync_poly_plot_df <- data.frame(c(seq(max_poly),seq(max_poly)),
                                 c(speed_dist_cv_error_10,speed_angle_cv_error_10),
                                 c(speed_dist_AIC_10,speed_angle_AIC_10),
                                 c(rep("Distance",max_poly),rep("Angle",max_poly)))

colnames(sync_poly_plot_df) <- c("Degree","Error","AIC","Predictor")

sync_poly_plot_df <- sync_poly_plot_df %>% group_by(Predictor) %>%
                                             mutate(minError = min(Error),minAIC = min(AIC)) %>%
                                             ungroup() %>%
                                             mutate(isMinEror = ifelse(Error == minError,3,1),isMinAIC = ifelse(AIC == minAIC,3,1))

ggplot(sync_poly_plot_df, aes(x = Degree, y = Error, color = Predictor))+
  geom_point(size = sync_poly_plot_df$isMinEror)+
  geom_line()+
  theme_light()+
  facet_wrap(~ Predictor, scales = "free") +
  scale_size(guide = "none")

ggplot(sync_poly_plot_df, aes(x = Degree, y = AIC, color = Predictor))+
  geom_point(size = sync_poly_plot_df$isMinAIC)+
  geom_line()+
  theme_light()+
  facet_wrap(~ Predictor, scales = "free") +
  scale_size(guide = "none")
```
Now let's use those polynomials to create the actual model

While they not be the lowest, the big improvements happen at 2 for Distance, and 2 for Angle

```{r}
sync_diff_glm <- glm(Sync ~ I(Distance^2)*I(Angle^2)*(Ablation+Flow+Darkness+Ablation:Flow+Darkness:Flow),
                        data = comp_data_sync_model)

summary(sync_diff_glm)
```

Now let's make some predictions

```{r}
sync_pred <- predict_df_da %>% mutate(Sync = predict(sync_diff_glm,predict_df_da))

ggplot(comp_data, aes(x = Distance, y = Sync))+
  geom_point(alpha = 0.1)+
  geom_smooth(method = lm, formula = y ~ poly(x, 2)) +
  facet_wrap(~ Flow + Ablation + Darkness) +
  theme_light()

ggplot(comp_data, aes(x = Angle, y = Sync))+
  geom_point(alpha = 0.1)+
  geom_smooth(method = glm, formula = y ~ poly(x, 2))+
  facet_wrap(~ Flow + Ablation + Darkness) +
  theme_light()

ggplot(sync_pred, aes(x = Distance, y = Sync))+
  geom_smooth()+
  scale_fill_viridis() +
  facet_wrap(~ Flow + Ablation + Darkness) +
  theme_light()

ggplot(sync_pred, aes(x = Angle, y = Sync))+
  geom_smooth()+
  scale_fill_viridis() +
  facet_wrap(~ Flow + Ablation + Darkness) +
  theme_light()
```