### 2022-05-18 Fish Data Analysis/Individual Fish.Rmd

---
title: "Individual Fish"
output: html_notebook
---

```{r}
library(tidyverse)
library(ggplot2)
library(car)
library(forcats)
library(ggpubr)
library(rstatix)
```

```{r}
indiv_data <- read.csv("Fish_Individual_Values.csv")
indiv_data <- na.omit(indiv_data)

indiv_data <- indiv_data %>% mutate(Flow = ifelse(Flow == "0", "Still Water", "Flowing Water (2 BL/s)")) %>%
                             mutate(Ablation = ifelse(Ablation == "N", "No Ablation", "Ablated")) %>%
                             mutate(Darkness = ifelse(Darkness == "N", "Light", "Dark")) %>%
                             mutate(Flow = factor(Flow), Ablation = factor(Ablation), Darkness = factor(Darkness)) %>%
                           mutate(Flow = fct_relevel(Flow, c("Still Water","Flowing Water (2 BL/s)"))) %>%
                           mutate(Ablation = fct_relevel(Ablation, c("No Ablation","Ablated"))) %>%
                           mutate(Darkness = fct_relevel(Darkness, c("Light","Dark"))) %>%
                           mutate(Flow_Ablation_Darkness = factor(paste(Flow,Ablation,Darkness,sep=", ")))

indiv_data_4_speed <- indiv_data %>% group_by(Year, Month, Day, Trial, Tailbeat_Num) %>%
                                     summarise(Mean_Speed = mean(Speed))
```

```{r}
m_speed <- aov(Speed ~ Flow + Ablation + Darkness + Flow:Ablation + Flow:Darkness, data = indiv_data)
Anova(m_speed)
TukeyHSD(m_speed)

ggplot(indiv_data,aes(x=Speed))+
  geom_histogram() +
  facet_wrap(~ Flow + Ablation + Darkness, scales="free") +
  xlim(0,6) +
  theme_light()

ggplot(indiv_data %>% filter(Speed < 5), aes(x = Flow, y = Speed, fill = interaction(Ablation,Darkness,sep=", ")))+
  geom_boxplot(outlier.shape = NA) +
  guides(fill = guide_legend(title = "Condition")) +
  scale_fill_manual(values=c("#EEEEEE", "#4CB944", "#00A7E1")) +
  ggtitle("Effect of Available Senses on Fish Speed") +
  xlab("") +
  #ylim(0,5) +
  ylab("Speed (BL/s)") +
  theme_light()

my_comparisons <- list( c("No Ablation, Light", "No Ablation, Dark"), c("No Ablation, Light", "Ablated, Light") )

ggplot(indiv_data %>% filter(Speed < 5), aes(x = interaction(Ablation,Darkness,sep=", "), y = Speed, fill = interaction(Ablation,Darkness,sep=", ")))+
  geom_boxplot(outlier.shape = NA) +
  guides(fill = guide_legend(title = "Condition")) +
  scale_fill_manual(values=c("#EEEEEE", "#4CB944", "#00A7E1")) +
  ggtitle("Effect of Available Senses on Fish Speed") +
  xlab("") +
  ylab("Speed (BL/s)") +
  ylim(0,6) +
  theme_light() +
  facet_wrap(~ Flow, strip.position = "bottom")+
  #stat_compare_means(aes(label = ..p.signif..), method = "t.test", ref.group = "Flow 0, No Ablation, Light")
  stat_compare_means(comparisons = my_comparisons, label.y = c(4.5, 5.5), label = "p.signif", hide.ns = TRUE) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```

```{r}
m_heading <- aov(Heading ~ Flow + Ablation + Darkness + Flow*Ablation + Flow*Darkness, data = indiv_data)
Anova(m_heading)
TukeyHSD(m_heading)

ggplot(indiv_data, aes(x = Flow, y = rad2deg(atan(sin(deg2rad(Heading))/cos(deg2rad(Heading)))), fill = interaction(Ablation,Darkness,sep=", ")))+
  geom_boxplot(outlier.shape = NA) +
  guides(fill = guide_legend(title = "Condition")) +
  scale_fill_manual(values=c("#EEEEEE", "#4CB944", "#00A7E1")) +
  ggtitle("Effect of Available Senses on Fish Heading") +
  xlab("") +
  ylab("Heading (Degrees)") +
  theme_light()
```

```{r}
m_tailbeats <- aov(TB_Frequency ~ Flow + Ablation + Darkness + Flow*Ablation + Flow*Darkness, data = indiv_data)
Anova(m_tailbeats)
TukeyHSD(m_tailbeats)

ggplot(indiv_data, aes(x = Flow, y = TB_Frequency, fill = interaction(Ablation,Darkness,sep=", ")))+
  geom_boxplot(outlier.shape = NA) +
  guides(fill = guide_legend(title = "Condition")) +
  scale_fill_manual(values=c("#EEEEEE", "#4CB944", "#00A7E1")) +
  ggtitle("Effect of Available Senses on Fish Tailbeat Frequency") +
  xlab("") +
  ylim(0,7) +
  ylab("Tailbeat Frequency (Beats/s)") +
  theme_light()

my_comparisons <- list(c("Still Water", "Flowing Water (2 BL/s)"))

ggplot(indiv_data, aes(x = Flow, y = TB_Frequency, fill = interaction(Ablation,Darkness,sep=", ")))+
  geom_boxplot(outlier.shape = NA) +
  guides(fill = guide_legend(title = "Condition")) +
  scale_fill_manual(values=c("#EEEEEE", "#4CB944", "#00A7E1")) +
  ggtitle("Effect of Available Senses on Fish Tailbeat Frequency") +
  xlab("") +
  ylab("Tailbeat Frequency (Beats/s)") +
  theme_light() +
  facet_wrap(~ interaction(Ablation,Darkness,sep=", "), strip.position = "bottom")+
   ylim(0,10) +
  #stat_compare_means(aes(label = ..p.signif..), method = "t.test", ref.group = "Flow 0, No Ablation, Light")
  stat_compare_means(comparisons = my_comparisons, label.y = c(9), label = "p.signif", hide.ns = TRUE) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

ggplot(indiv_data, aes(x = Flow, y = TB_Frequency))+
  geom_boxplot(outlier.shape = NA) +
  guides(fill = guide_legend(title = "Condition")) +
  scale_fill_manual(values=c("#EEEEEE", "#4CB944", "#00A7E1")) +
  ggtitle("Effect of Available Senses on Fish Tailbeat Frequency") +
  xlab("") +
  ylab("Tailbeat Frequency (Beats/s)") +
  theme_light() +
  facet_wrap(~ interaction(Ablation,Darkness,sep=", "), strip.position = "bottom")+
   ylim(0,10) +
  #stat_compare_means(aes(label = ..p.signif..), method = "t.test", ref.group = "Flow 0, No Ablation, Light")
  stat_compare_means(comparisons = my_comparisons, label.y = c(9), label = "p.signif", hide.ns = TRUE) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```