### 2022-05-18 Fish Data Analysis/comp_fish_simple_models.Rmd

---
title: "Fish Comparisons Simple Models"
output: html_notebook
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(car)
library(viridis)
library(boot)
library(mgcv)
library(forcats)
library(ggpubr)
library(mgcViz)
```

Alright, so from all of this what do I know?

Speed Differences:
  Flow increases speed differences
  Ablation decreases speed differences
  Darkness increases speed differences
  Distance increases speed differences
  Angles around 90 degrees have a bigger difference in speed

Heading Differences:
  Fish are more aligned when they are around 90 degrees, slightly less in ablated fish

Synchronization:
  Darkness increases synchronization

NND:
  Darkness increases NND
  Flow decreases NND

Nearest Neighbor Bearing:
  In darkness fish are more in front or behind
  In flow fish are more in positioned alongside each other

Sets up the functions that will be used later 
```{r}
rad2deg <- function(rad) {(rad * 180) / (pi)}
deg2rad <- function(deg) {(deg * pi) / (180)}
round_any <- function(x, accuracy, f=round){f(x/ accuracy) * accuracy}
ang_mean <- function(x){rad2deg(atan2(mean(sin(deg2rad(x))),mean(cos(deg2rad(x)))))}

fold_angle_0_360_to_0_180 <- function(x){abs(abs(x-180)-180)}

fold_angle_neg_180_180_to_neg_90_90 <- function(x){ifelse(x > 90,90-abs(90-x),ifelse(x < -90, -90+abs(-90-x), x))}

min_n <- function(x,n){sort(x)[1:n]}
max_n <- function(x,n){sort(x,decreasing = TRUE)[1:n]}
```

Reads in the data and alters it as needed 
```{r}
comp_data <- read.csv("Fish_Comp_Values.csv")
comp_data <- na.omit(comp_data)

comp_data <- comp_data %>% mutate(Flow = ifelse(Flow == "0", "Still Water", "Flowing Water (2 BL/s)")) %>%
                           mutate(Ablation = ifelse(Ablation == "N", "No Ablation", "Ablated")) %>%
                           mutate(Darkness = ifelse(Darkness == "N", "Light", "Dark")) %>%
                           filter(Distance <= 4) %>%
                           filter(abs(Speed_Diff) <= 6) %>% 
                           #mutate(Speed_Diff = abs(Speed_Diff)) %>%
                           mutate(rad_Heading = deg2rad(Heading_Diff)) %>%
                           mutate(sin_Heading = sin(rad_Heading), cos_Heading = cos(rad_Heading)) %>%
                           #mutate(fold_heading_diff = abs(fold_angle_neg_180_180_to_neg_90_90(Heading_Diff))) %>%
                           mutate(Is_Aligned = ifelse(Heading_Diff < 30, 1, 0)) %>%
                           mutate(Is_Reversed = ifelse(Heading_Diff > 150, 1, 0)) %>%
                           separate(Fish, c("Fish1", "Fish2"),sep="x") %>%
                           mutate(Flow = factor(Flow), Ablation = factor(Ablation), Darkness = factor(Darkness)) %>%
                           mutate(Flow = fct_relevel(Flow, c("Still Water","Flowing Water (2 BL/s)"))) %>%
                           mutate(Ablation = fct_relevel(Ablation, c("No Ablation","Ablated"))) %>%
                           mutate(Darkness = fct_relevel(Darkness, c("Light","Dark"))) %>%
                           mutate(rad_Angle = deg2rad(Angle)) %>%
                           mutate(sin_Angle = sin(rad_Angle), cos_Angle = cos(rad_Angle)) %>%
                           mutate(Flow_Ablation = factor(paste(Flow,Ablation,sep=", "))) %>%
                           mutate(Flow_Darkness = factor(paste(Flow,Darkness,sep=", "))) %>%
                           mutate(Flow_Ablation_Darkness = factor(paste(Flow,Ablation,Darkness,sep=", ")))

                           #mutate(Angle = fold_angle_0_360_to_0_180(Angle))

sum_comp_data <- comp_data %>% mutate(X_Distance = round_any(X_Distance,0.25), 
                                      Y_Distance = round_any(abs(Y_Distance),0.25)) %>%
                               group_by(Flow,Ablation,Darkness,X_Distance,Y_Distance) %>%
                               summarise(Speed_Diff = mean(Speed_Diff),
                                         Heading_Diff = ang_mean(Heading_Diff),
                                         Sync = mean(Sync),
                                         #Fold_Heading_Diff = mean(fold_heading_diff),
                                         Is_Aligned = mean(Is_Aligned),
                                         Is_Reversed = mean(Is_Reversed))

comp_data_switch_fish <- comp_data %>% mutate(Fish3 = Fish1) %>%
                                       mutate(Fish1 = Fish2) %>%
                                       mutate(Fish2 = Fish3) %>%
                                       select(-c(Fish3))

comp_data_NND <- bind_rows(comp_data, comp_data_switch_fish) %>% group_by(Year,Month,Day,Trial,Fish1,Tailbeat_Num) %>%
                                                                 filter(Distance == min(Distance)) %>%
                                                                 ungroup()

comp_data_near_3 <- bind_rows(comp_data, comp_data_switch_fish) %>% group_by(Year,Month,Day,Trial,Fish1,Tailbeat_Num) %>%
                                                                 slice_min(Distance, n = 3) %>%
                                                                 ungroup() %>%
                                                                 distinct(Distance, .keep_all= TRUE)
```

#Basic Stats

```{r}
speed_anova <- aov(Speed_Diff ~ Flow + Ablation + Darkness + Flow:Ablation + Flow:Darkness, data = comp_data_NND)

Anova(speed_anova)

heading_anova <- aov(Heading_Diff ~ Flow + Ablation + Darkness + Flow:Ablation + Flow:Darkness, data = comp_data_NND)

Anova(heading_anova)

sync_anova <- aov(Sync ~ Flow + Ablation + Darkness + Flow:Ablation + Flow:Darkness, data = comp_data_NND)

Anova(sync_anova)

dist_anova <- aov(Distance ~ Flow + Ablation + Darkness + Flow:Ablation + Flow:Darkness, data = comp_data_NND)

Anova(dist_anova)

angle_anova <- aov(Angle ~ Flow + Ablation + Darkness + Flow:Ablation + Flow:Darkness, data = comp_data_NND)

Anova(angle_anova)
```

```{r}
ggplot(comp_data_near_3, aes(x = Flow, y = Speed_Diff, fill = interaction(Ablation,Darkness,sep=", ")))+
  geom_boxplot(outlier.shape = NA) +
  guides(fill = guide_legend(title = "Condition")) +
  scale_fill_manual(values=c("#EEEEEE", "#4CB944", "#00A7E1")) +
  ggtitle("Effect of Available Senses on Speed Difference") +
  xlab("") +
  ylab("Speed Difference (BL/s)") +
  #ylim(-0.1,2.5) +
  theme_light()

#rad2deg(atan(sin_Heading/cos_Heading))

ggplot(comp_data_near_3, aes(x = Flow, y = Heading_Diff, fill = interaction(Ablation,Darkness,sep=", ")))+
  geom_violin(outlier.shape = NA) +
  guides(fill = guide_legend(title = "Condition")) +
  scale_fill_manual(values=c("#EEEEEE", "#4CB944", "#00A7E1")) +
  ggtitle("Effect of Available Senses on Heading Difference") +
  xlab("") +
  ylab("Heading Difference (Degrees)") +
  #scale_y_continuous(breaks = c(-90,-60,-30,0,30,60,90)) +
  theme_light()

ggplot(comp_data_near_3, aes(x = Flow, y = Sync, fill = interaction(Ablation,Darkness,sep=", ")))+
  geom_boxplot(outlier.shape = NA) +
  guides(fill = guide_legend(title = "Condition")) +
  scale_fill_manual(values=c("#EEEEEE", "#4CB944", "#00A7E1")) +
  ggtitle("Effect of Available Senses on Tailbeat Synchonization") +
  xlab("") +
  ylab("Syncronization") +
  theme_light()

ggplot(comp_data_near_3, aes(x = Flow, y = Distance, fill = interaction(Ablation,Darkness,sep=", ")))+
  geom_boxplot(outlier.shape = NA) +
  guides(fill = guide_legend(title = "Condition")) +
  scale_fill_manual(values=c("#EEEEEE", "#4CB944", "#00A7E1")) +
  ggtitle("Effect of Available Senses on Nearest Neighbor Distance (NND)") +
  xlab("") +
  ylab("NND (BL)") +
  theme_light()

#rad2deg(atan2(sin_Angle,cos_Angle))

ggplot(comp_data_near_3, aes(x = Flow, y = Angle, fill = interaction(Ablation,Darkness,sep=", ")))+
  geom_violin(outlier.shape = NA) +
  guides(fill = guide_legend(title = "Condition")) +
  scale_fill_manual(values=c("#EEEEEE", "#4CB944", "#00A7E1")) +
  ggtitle("Effect of Available Senses on Nearest Neighbor Bearing") +
  xlab("") +
  ylab("Nearest Neighbor Bearing (Degrees)") +
  #scale_y_continuous(breaks = c(-90,-60,-30,0,30,60,90)) +
  theme_light()

```
 Just nearest Fish Speed Diff

```{r}
comp_data_near_1 <- bind_rows(comp_data, comp_data_switch_fish) %>% group_by(Year,Month,Day,Trial,Fish1,Tailbeat_Num) %>%
                                                                 slice_max(Distance, n = 1) %>%
                                                                 ungroup() %>%
                                                                 distinct(Distance, .keep_all= TRUE)

speed_anova_nn <- aov(Speed_Diff ~ Flow + Ablation + Darkness + Flow:Ablation + Flow:Darkness, data = comp_data_near_1)

Anova(speed_anova_nn)

my_comparisons <- list( c("No Ablation, Light", "No Ablation, Dark"))

ggplot(comp_data_near_1, aes(x = interaction(Ablation,Darkness,sep=", "), y = Speed_Diff, fill = interaction(Ablation,Darkness,sep=", ")))+
  geom_violin(outlier.shape = NA) +
  guides(fill = guide_legend(title = "Condition")) +
  scale_fill_manual(values=c("#EEEEEE", "#4CB944", "#00A7E1")) +
  ggtitle("Effect of Available Senses on Nearest Speed Diff") +
  xlab("") +
  ylab("Speed Difference (BL/s)") +
  theme_light() +
  facet_wrap(~ Flow, strip.position = "bottom")+
  #stat_compare_means(aes(label = ..p.signif..), method = "t.test", ref.group = "Flow 0, No Ablation, Light")
  #stat_compare_means(comparisons = my_comparisons, label.y = c(7.5,8.5), label = "p.signif", hide.ns = TRUE) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```

Trying ggpubr

```{r}
my_comparisons <- list( c("No Ablation, Light", "No Ablation, Dark") )

ggplot(comp_data_near_3, aes(x = interaction(Ablation,Darkness,sep=", "), y = Distance, fill = interaction(Ablation,Darkness,sep=", ")))+
  geom_boxplot(outlier.shape = NA) +
  guides(fill = guide_legend(title = "Condition")) +
  scale_fill_manual(values=c("#EEEEEE", "#4CB944", "#00A7E1")) +
  ggtitle("Effect of Available Senses on Nearest Neighbor Distance (NND)") +
  xlab("") +
  ylab("NND (BL)") +
  theme_light() +
  facet_wrap(~ Flow)+
  #stat_compare_means(aes(label = ..p.signif..), method = "t.test", ref.group = "Flow 0, No Ablation, Light")
  stat_compare_means(comparisons = my_comparisons, label.y = c(4.5), label = "p.signif", hide.ns = TRUE) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```

So we need to make all the possible GLMs for these variables and then compare them in order to see which model does best in order to see what actually has effects
```{r}
# speed_gam_d <- gam(Speed_Diff ~ s(Distance), data = comp_data_near_3)
# speed_gam_d_f <- gam(Speed_Diff ~ s(Distance,by=Flow), data = comp_data_near_3)
# speed_gam_d_a <- gam(Speed_Diff ~ s(Distance,by=Ablation), data = comp_data_near_3)
# speed_gam_d_d <- gam(Speed_Diff ~ s(Distance,by=Darkness), data = comp_data_near_3)
# speed_gam_d_fa <- gam(Speed_Diff ~ s(Distance,by=Flow_Ablation), data = comp_data_near_3)
# speed_gam_d_fd <- gam(Speed_Diff ~ s(Distance,by=Flow_Darkness), data = comp_data_near_3)
# speed_gam_d_fda <- gam(Speed_Diff ~ s(Distance,by=Flow_Ablation_Darkness), data = comp_data_near_3)
# 
# speed_gam_ca <- gam(Speed_Diff ~ s(cos_Angle), data = comp_data_near_3)
# speed_gam_ca_f <- gam(Speed_Diff ~ s(cos_Angle,by=Flow), data = comp_data_near_3)
# speed_gam_ca_a <- gam(Speed_Diff ~ s(cos_Angle,by=Ablation), data = comp_data_near_3)
# speed_gam_ca_d <- gam(Speed_Diff ~ s(cos_Angle,by=Darkness), data = comp_data_near_3)
# speed_gam_ca_fa <- gam(Speed_Diff ~ s(cos_Angle,by=Flow_Ablation), data = comp_data_near_3)
# speed_gam_ca_fd <- gam(Speed_Diff ~ s(cos_Angle,by=Flow_Darkness), data = comp_data_near_3)
# speed_gam_ca_fda <- gam(Speed_Diff ~ s(cos_Angle,by=Flow_Ablation_Darkness), data = comp_data_near_3)
# 
# speed_gam_sa <- gam(Speed_Diff ~ s(sin_Angle), data = comp_data_near_3)
# speed_gam_sa_f <- gam(Speed_Diff ~ s(sin_Angle,by=Flow), data = comp_data_near_3)
# speed_gam_sa_a <- gam(Speed_Diff ~ s(sin_Angle,by=Ablation), data = comp_data_near_3)
# speed_gam_sa_d <- gam(Speed_Diff ~ s(sin_Angle,by=Darkness), data = comp_data_near_3)
# speed_gam_sa_fa <- gam(Speed_Diff ~ s(sin_Angle,by=Flow_Ablation), data = comp_data_near_3)
# speed_gam_sa_fd <- gam(Speed_Diff ~ s(sin_Angle,by=Flow_Darkness), data = comp_data_near_3)
# speed_gam_sa_fda <- gam(Speed_Diff ~ s(sin_Angle,by=Flow_Ablation_Darkness), data = comp_data_near_3)
# 
# speed_gam_dca <- gam(Speed_Diff ~ s(Distance,cos_Angle), data = comp_data_near_3)
# speed_gam_dca_f <- gam(Speed_Diff ~ s(Distance,cos_Angle,by=Flow), data = comp_data_near_3)
# speed_gam_dca_a <- gam(Speed_Diff ~ s(Distance,cos_Angle,by=Ablation), data = comp_data_near_3)
# speed_gam_dca_d <- gam(Speed_Diff ~ s(Distance,cos_Angle,by=Darkness), data = comp_data_near_3)
# speed_gam_dca_fa <- gam(Speed_Diff ~ s(Distance,cos_Angle,by=Flow_Ablation), data = comp_data_near_3)
# speed_gam_dca_fd <- gam(Speed_Diff ~ s(Distance,cos_Angle,by=Flow_Darkness), data = comp_data_near_3)
# speed_gam_dca_fda <- gam(Speed_Diff ~ s(Distance,cos_Angle,by=Flow_Ablation_Darkness), data = comp_data_near_3)
# 
# speed_gam_dsa <- gam(Speed_Diff ~ s(Distance,sin_Angle), data = comp_data_near_3)
# speed_gam_dsa_f <- gam(Speed_Diff ~ s(Distance,sin_Angle,by=Flow), data = comp_data_near_3)
# speed_gam_dsa_a <- gam(Speed_Diff ~ s(Distance,sin_Angle,by=Ablation), data = comp_data_near_3)
# speed_gam_dsa_d <- gam(Speed_Diff ~ s(Distance,sin_Angle,by=Darkness), data = comp_data_near_3)
# speed_gam_dsa_fa <- gam(Speed_Diff ~ s(Distance,sin_Angle,by=Flow_Ablation), data = comp_data_near_3)
# speed_gam_dsa_fd <- gam(Speed_Diff ~ s(Distance,sin_Angle,by=Flow_Darkness), data = comp_data_near_3)
# speed_gam_dsa_fda <- gam(Speed_Diff ~ s(Distance,sin_Angle,by=Flow_Ablation_Darkness), data = comp_data_near_3)
# 
# speed_gam_dcasa <- gam(Speed_Diff ~ s(Distance,cos_Angle,sin_Angle), data = comp_data_near_3)
# speed_gam_dcasa_f <- gam(Speed_Diff ~ s(Distance,cos_Angle,sin_Angle,by=Flow), data = comp_data_near_3)
# speed_gam_dcasa_a <- gam(Speed_Diff ~ s(Distance,cos_Angle,sin_Angle,by=Ablation), data = comp_data_near_3)
# speed_gam_dcasa_d <- gam(Speed_Diff ~ s(Distance,cos_Angle,sin_Angle,by=Darkness), data = comp_data_near_3)
# speed_gam_dcasa_fa <- gam(Speed_Diff ~ s(Distance,cos_Angle,sin_Angle,by=Flow_Ablation), data = comp_data_near_3)
# speed_gam_dcasa_fd <- gam(Speed_Diff ~ s(Distance,cos_Angle,sin_Angle,by=Flow_Darkness), data = comp_data_near_3)
# speed_gam_dcasa_fda <- gam(Speed_Diff ~ s(Distance,cos_Angle,sin_Angle,by=Flow_Ablation_Darkness), data = comp_data_near_3)

speed_AIC <- AIC(speed_gam_d,speed_gam_d_f,speed_gam_d_a,speed_gam_d_d,speed_gam_d_fa,speed_gam_d_fd,speed_gam_d_fda,
                 speed_gam_ca,speed_gam_ca_f,speed_gam_ca_a,speed_gam_ca_d,speed_gam_ca_fa,speed_gam_ca_fd,speed_gam_ca_fda,
                 speed_gam_sa,speed_gam_sa_f,speed_gam_sa_a,speed_gam_sa_d,speed_gam_sa_fa,speed_gam_sa_fd,speed_gam_sa_fda,
                 speed_gam_dca,speed_gam_dca_f,speed_gam_dca_a,speed_gam_dca_d,speed_gam_dca_fa,speed_gam_dca_fd,speed_gam_dca_fda,
                 speed_gam_dsa,speed_gam_dsa_f,speed_gam_dsa_a,speed_gam_dsa_d,speed_gam_dsa_fa,speed_gam_dsa_fd,speed_gam_dsa_fda,
                 speed_gam_dcasa,speed_gam_dcasa_f,speed_gam_dcasa_a,speed_gam_dcasa_d,speed_gam_dcasa_fa,speed_gam_dcasa_fd,speed_gam_dcasa_fda)

speed_AIC <- tibble::rownames_to_column(speed_AIC, "Model_Name")

speed_AIC <- speed_AIC %>% mutate(Model_Name = fct_reorder(Model_Name, -AIC)) %>% arrange(AIC)

ggplot(speed_AIC, aes(x = Model_Name, y = AIC)) +
  geom_point(color = "blue") +
  xlab("") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme_light()+
  scale_y_reverse() +
  coord_flip()

speed_AIC
```

How we do that for the cosine of the headings

```{r}
# cos_heading_gam_d <- gam(cos_Heading ~ s(Distance), data = comp_data_near_3)
# cos_heading_gam_d_f <- gam(cos_Heading ~ s(Distance,by=Flow), data = comp_data_near_3)
# cos_heading_gam_d_a <- gam(cos_Heading ~ s(Distance,by=Ablation), data = comp_data_near_3)
# cos_heading_gam_d_d <- gam(cos_Heading ~ s(Distance,by=Darkness), data = comp_data_near_3)
# cos_heading_gam_d_fa <- gam(cos_Heading ~ s(Distance,by=Flow_Ablation), data = comp_data_near_3)
# cos_heading_gam_d_fd <- gam(cos_Heading ~ s(Distance,by=Flow_Darkness), data = comp_data_near_3)
# cos_heading_gam_d_fda <- gam(cos_Heading ~ s(Distance,by=Flow_Ablation_Darkness), data = comp_data_near_3)
# 
# cos_heading_gam_ca <- gam(cos_Heading ~ s(cos_Angle), data = comp_data_near_3)
# cos_heading_gam_ca_f <- gam(cos_Heading ~ s(cos_Angle,by=Flow), data = comp_data_near_3)
# cos_heading_gam_ca_a <- gam(cos_Heading ~ s(cos_Angle,by=Ablation), data = comp_data_near_3)
# cos_heading_gam_ca_d <- gam(cos_Heading ~ s(cos_Angle,by=Darkness), data = comp_data_near_3)
# cos_heading_gam_ca_fa <- gam(cos_Heading ~ s(cos_Angle,by=Flow_Ablation), data = comp_data_near_3)
# cos_heading_gam_ca_fd <- gam(cos_Heading ~ s(cos_Angle,by=Flow_Darkness), data = comp_data_near_3)
# cos_heading_gam_ca_fda <- gam(cos_Heading ~ s(cos_Angle,by=Flow_Ablation_Darkness), data = comp_data_near_3)
# 
# cos_heading_gam_sa <- gam(cos_Heading ~ s(sin_Angle), data = comp_data_near_3)
# cos_heading_gam_sa_f <- gam(cos_Heading ~ s(sin_Angle,by=Flow), data = comp_data_near_3)
# cos_heading_gam_sa_a <- gam(cos_Heading ~ s(sin_Angle,by=Ablation), data = comp_data_near_3)
# cos_heading_gam_sa_d <- gam(cos_Heading ~ s(sin_Angle,by=Darkness), data = comp_data_near_3)
# cos_heading_gam_sa_fa <- gam(cos_Heading ~ s(sin_Angle,by=Flow_Ablation), data = comp_data_near_3)
# cos_heading_gam_sa_fd <- gam(cos_Heading ~ s(sin_Angle,by=Flow_Darkness), data = comp_data_near_3)
# cos_heading_gam_sa_fda <- gam(cos_Heading ~ s(sin_Angle,by=Flow_Ablation_Darkness), data = comp_data_near_3)
# 
# cos_heading_gam_dca <- gam(cos_Heading ~ s(Distance,cos_Angle), data = comp_data_near_3)
# cos_heading_gam_dca_f <- gam(cos_Heading ~ s(Distance,cos_Angle,by=Flow), data = comp_data_near_3)
# cos_heading_gam_dca_a <- gam(cos_Heading ~ s(Distance,cos_Angle,by=Ablation), data = comp_data_near_3)
# cos_heading_gam_dca_d <- gam(cos_Heading ~ s(Distance,cos_Angle,by=Darkness), data = comp_data_near_3)
# cos_heading_gam_dca_fa <- gam(cos_Heading ~ s(Distance,cos_Angle,by=Flow_Ablation), data = comp_data_near_3)
# cos_heading_gam_dca_fd <- gam(cos_Heading ~ s(Distance,cos_Angle,by=Flow_Darkness), data = comp_data_near_3)
# cos_heading_gam_dca_fda <- gam(cos_Heading ~ s(Distance,cos_Angle,by=Flow_Ablation_Darkness), data = comp_data_near_3)
# 
# cos_heading_gam_dsa <- gam(cos_Heading ~ s(Distance,sin_Angle), data = comp_data_near_3)
# cos_heading_gam_dsa_f <- gam(cos_Heading ~ s(Distance,sin_Angle,by=Flow), data = comp_data_near_3)
# cos_heading_gam_dsa_a <- gam(cos_Heading ~ s(Distance,sin_Angle,by=Ablation), data = comp_data_near_3)
# cos_heading_gam_dsa_d <- gam(cos_Heading ~ s(Distance,sin_Angle,by=Darkness), data = comp_data_near_3)
# cos_heading_gam_dsa_fa <- gam(cos_Heading ~ s(Distance,sin_Angle,by=Flow_Ablation), data = comp_data_near_3)
# cos_heading_gam_dsa_fd <- gam(cos_Heading ~ s(Distance,sin_Angle,by=Flow_Darkness), data = comp_data_near_3)
# cos_heading_gam_dsa_fda <- gam(cos_Heading ~ s(Distance,sin_Angle,by=Flow_Ablation_Darkness), data = comp_data_near_3)
# 
# cos_heading_gam_dcasa <- gam(cos_Heading ~ s(Distance,cos_Angle,sin_Angle), data = comp_data_near_3)
# cos_heading_gam_dcasa_f <- gam(cos_Heading ~ s(Distance,cos_Angle,sin_Angle,by=Flow), data = comp_data_near_3)
# cos_heading_gam_dcasa_a <- gam(cos_Heading ~ s(Distance,cos_Angle,sin_Angle,by=Ablation), data = comp_data_near_3)
# cos_heading_gam_dcasa_d <- gam(cos_Heading ~ s(Distance,cos_Angle,sin_Angle,by=Darkness), data = comp_data_near_3)
# cos_heading_gam_dcasa_fa <- gam(cos_Heading ~ s(Distance,cos_Angle,sin_Angle,by=Flow_Ablation), data = comp_data_near_3)
# cos_heading_gam_dcasa_fd <- gam(cos_Heading ~ s(Distance,cos_Angle,sin_Angle,by=Flow_Darkness), data = comp_data_near_3)
# cos_heading_gam_dcasa_fda <- gam(cos_Heading ~ s(Distance,cos_Angle,sin_Angle,by=Flow_Ablation_Darkness), data = comp_data_near_3)

cos_heading_AIC <- AIC(cos_heading_gam_d,cos_heading_gam_d_f,cos_heading_gam_d_a,cos_heading_gam_d_d,cos_heading_gam_d_fa,cos_heading_gam_d_fd,cos_heading_gam_d_fda,
                 cos_heading_gam_ca,cos_heading_gam_ca_f,cos_heading_gam_ca_a,cos_heading_gam_ca_d,cos_heading_gam_ca_fa,cos_heading_gam_ca_fd,cos_heading_gam_ca_fda,
                 cos_heading_gam_sa,cos_heading_gam_sa_f,cos_heading_gam_sa_a,cos_heading_gam_sa_d,cos_heading_gam_sa_fa,cos_heading_gam_sa_fd,cos_heading_gam_sa_fda,
                 cos_heading_gam_dca,cos_heading_gam_dca_f,cos_heading_gam_dca_a,cos_heading_gam_dca_d,cos_heading_gam_dca_fa,cos_heading_gam_dca_fd,cos_heading_gam_dca_fda,
                 cos_heading_gam_dsa,cos_heading_gam_dsa_f,cos_heading_gam_dsa_a,cos_heading_gam_dsa_d,cos_heading_gam_dsa_fa,cos_heading_gam_dsa_fd,cos_heading_gam_dsa_fda,
                 cos_heading_gam_dcasa,cos_heading_gam_dcasa_f,cos_heading_gam_dcasa_a,cos_heading_gam_dcasa_d,cos_heading_gam_dcasa_fa,cos_heading_gam_dcasa_fd,cos_heading_gam_dcasa_fda)

cos_heading_AIC <- tibble::rownames_to_column(cos_heading_AIC, "Model_Name")

cos_heading_AIC <- cos_heading_AIC %>% mutate(Model_Name = fct_reorder(Model_Name, -AIC)) %>% arrange(AIC)

ggplot(cos_heading_AIC %>% arrange(AIC), aes(x = Model_Name, y = AIC)) +
  geom_point(color = "blue") +
  xlab("") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme_light() +
  scale_y_reverse() +
  coord_flip()

cos_heading_AIC
```

Now with the sin of the Headings
```{r}
# sin_heading_gam_d <- gam(sin_Heading ~ s(Distance), data = comp_data_near_3)
# sin_heading_gam_d_f <- gam(sin_Heading ~ s(Distance,by=Flow), data = comp_data_near_3)
# sin_heading_gam_d_a <- gam(sin_Heading ~ s(Distance,by=Ablation), data = comp_data_near_3)
# sin_heading_gam_d_d <- gam(sin_Heading ~ s(Distance,by=Darkness), data = comp_data_near_3)
# sin_heading_gam_d_fa <- gam(sin_Heading ~ s(Distance,by=Flow_Ablation), data = comp_data_near_3)
# sin_heading_gam_d_fd <- gam(sin_Heading ~ s(Distance,by=Flow_Darkness), data = comp_data_near_3)
# sin_heading_gam_d_fda <- gam(sin_Heading ~ s(Distance,by=Flow_Ablation_Darkness), data = comp_data_near_3)
# 
# sin_heading_gam_ca <- gam(sin_Heading ~ s(cos_Angle), data = comp_data_near_3)
# sin_heading_gam_ca_f <- gam(sin_Heading ~ s(cos_Angle,by=Flow), data = comp_data_near_3)
# sin_heading_gam_ca_a <- gam(sin_Heading ~ s(cos_Angle,by=Ablation), data = comp_data_near_3)
# sin_heading_gam_ca_d <- gam(sin_Heading ~ s(cos_Angle,by=Darkness), data = comp_data_near_3)
# sin_heading_gam_ca_fa <- gam(sin_Heading ~ s(cos_Angle,by=Flow_Ablation), data = comp_data_near_3)
# sin_heading_gam_ca_fd <- gam(sin_Heading ~ s(cos_Angle,by=Flow_Darkness), data = comp_data_near_3)
# sin_heading_gam_ca_fda <- gam(sin_Heading ~ s(cos_Angle,by=Flow_Ablation_Darkness), data = comp_data_near_3)
# 
# sin_heading_gam_sa <- gam(sin_Heading ~ s(sin_Angle), data = comp_data_near_3)
# sin_heading_gam_sa_f <- gam(sin_Heading ~ s(sin_Angle,by=Flow), data = comp_data_near_3)
# sin_heading_gam_sa_a <- gam(sin_Heading ~ s(sin_Angle,by=Ablation), data = comp_data_near_3)
# sin_heading_gam_sa_d <- gam(sin_Heading ~ s(sin_Angle,by=Darkness), data = comp_data_near_3)
# sin_heading_gam_sa_fa <- gam(sin_Heading ~ s(sin_Angle,by=Flow_Ablation), data = comp_data_near_3)
# sin_heading_gam_sa_fd <- gam(sin_Heading ~ s(sin_Angle,by=Flow_Darkness), data = comp_data_near_3)
# sin_heading_gam_sa_fda <- gam(sin_Heading ~ s(sin_Angle,by=Flow_Ablation_Darkness), data = comp_data_near_3)
# 
# sin_heading_gam_dca <- gam(sin_Heading ~ s(Distance,cos_Angle), data = comp_data_near_3)
# sin_heading_gam_dca_f <- gam(sin_Heading ~ s(Distance,cos_Angle,by=Flow), data = comp_data_near_3)
# sin_heading_gam_dca_a <- gam(sin_Heading ~ s(Distance,cos_Angle,by=Ablation), data = comp_data_near_3)
# sin_heading_gam_dca_d <- gam(sin_Heading ~ s(Distance,cos_Angle,by=Darkness), data = comp_data_near_3)
# sin_heading_gam_dca_fa <- gam(sin_Heading ~ s(Distance,cos_Angle,by=Flow_Ablation), data = comp_data_near_3)
# sin_heading_gam_dca_fd <- gam(sin_Heading ~ s(Distance,cos_Angle,by=Flow_Darkness), data = comp_data_near_3)
# sin_heading_gam_dca_fda <- gam(sin_Heading ~ s(Distance,cos_Angle,by=Flow_Ablation_Darkness), data = comp_data_near_3)
# 
# sin_heading_gam_dsa <- gam(sin_Heading ~ s(Distance,sin_Angle), data = comp_data_near_3)
# sin_heading_gam_dsa_f <- gam(sin_Heading ~ s(Distance,sin_Angle,by=Flow), data = comp_data_near_3)
# sin_heading_gam_dsa_a <- gam(sin_Heading ~ s(Distance,sin_Angle,by=Ablation), data = comp_data_near_3)
# sin_heading_gam_dsa_d <- gam(sin_Heading ~ s(Distance,sin_Angle,by=Darkness), data = comp_data_near_3)
# sin_heading_gam_dsa_fa <- gam(sin_Heading ~ s(Distance,sin_Angle,by=Flow_Ablation), data = comp_data_near_3)
# sin_heading_gam_dsa_fd <- gam(sin_Heading ~ s(Distance,sin_Angle,by=Flow_Darkness), data = comp_data_near_3)
# sin_heading_gam_dsa_fda <- gam(sin_Heading ~ s(Distance,sin_Angle,by=Flow_Ablation_Darkness), data = comp_data_near_3)
# 
# sin_heading_gam_dcasa <- gam(sin_Heading ~ s(Distance,cos_Angle,sin_Angle), data = comp_data_near_3)
# sin_heading_gam_dcasa_f <- gam(sin_Heading ~ s(Distance,cos_Angle,sin_Angle,by=Flow), data = comp_data_near_3)
# sin_heading_gam_dcasa_a <- gam(sin_Heading ~ s(Distance,cos_Angle,sin_Angle,by=Ablation), data = comp_data_near_3)
# sin_heading_gam_dcasa_d <- gam(sin_Heading ~ s(Distance,cos_Angle,sin_Angle,by=Darkness), data = comp_data_near_3)
# sin_heading_gam_dcasa_fa <- gam(sin_Heading ~ s(Distance,cos_Angle,sin_Angle,by=Flow_Ablation), data = comp_data_near_3)
# sin_heading_gam_dcasa_fd <- gam(sin_Heading ~ s(Distance,cos_Angle,sin_Angle,by=Flow_Darkness), data = comp_data_near_3)
# sin_heading_gam_dcasa_fda <- gam(sin_Heading ~ s(Distance,cos_Angle,sin_Angle,by=Flow_Ablation_Darkness), data = comp_data_near_3)

sin_Heading_AIC <- AIC(sin_heading_gam_d,sin_heading_gam_d_f,sin_heading_gam_d_a,sin_heading_gam_d_d,sin_heading_gam_d_fa,sin_heading_gam_d_fd,sin_heading_gam_d_fda,
                 sin_heading_gam_ca,sin_heading_gam_ca_f,sin_heading_gam_ca_a,sin_heading_gam_ca_d,sin_heading_gam_ca_fa,sin_heading_gam_ca_fd,sin_heading_gam_ca_fda,
                 sin_heading_gam_sa,sin_heading_gam_sa_f,sin_heading_gam_sa_a,sin_heading_gam_sa_d,sin_heading_gam_sa_fa,sin_heading_gam_sa_fd,sin_heading_gam_sa_fda,
                 sin_heading_gam_dca,sin_heading_gam_dca_f,sin_heading_gam_dca_a,sin_heading_gam_dca_d,sin_heading_gam_dca_fa,sin_heading_gam_dca_fd,sin_heading_gam_dca_fda,
                 sin_heading_gam_dsa,sin_heading_gam_dsa_f,sin_heading_gam_dsa_a,sin_heading_gam_dsa_d,sin_heading_gam_dsa_fa,sin_heading_gam_dsa_fd,sin_heading_gam_dsa_fda,
                 sin_heading_gam_dcasa,sin_heading_gam_dcasa_f,sin_heading_gam_dcasa_a,sin_heading_gam_dcasa_d,sin_heading_gam_dcasa_fa,sin_heading_gam_dcasa_fd,sin_heading_gam_dcasa_fda)

sin_Heading_AIC <- tibble::rownames_to_column(sin_Heading_AIC, "Model_Name")

sin_Heading_AIC <- sin_Heading_AIC %>% mutate(Model_Name = fct_reorder(Model_Name, -AIC)) %>% arrange(AIC)

ggplot(sin_Heading_AIC %>% arrange(AIC), aes(x = Model_Name, y = AIC)) +
  geom_point(color = "blue") +
  xlab("") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme_light()+
  scale_y_reverse() +
  coord_flip()

sin_Heading_AIC
```

And now with Sync

```{r}
# sync_gam_d <- gam(Sync ~ s(Distance), data = comp_data_near_3)
# sync_gam_d_f <- gam(Sync ~ s(Distance,by=Flow), data = comp_data_near_3)
# sync_gam_d_a <- gam(Sync ~ s(Distance,by=Ablation), data = comp_data_near_3)
# sync_gam_d_d <- gam(Sync ~ s(Distance,by=Darkness), data = comp_data_near_3)
# sync_gam_d_fa <- gam(Sync ~ s(Distance,by=Flow_Ablation), data = comp_data_near_3)
# sync_gam_d_fd <- gam(Sync ~ s(Distance,by=Flow_Darkness), data = comp_data_near_3)
# sync_gam_d_fda <- gam(Sync ~ s(Distance,by=Flow_Ablation_Darkness), data = comp_data_near_3)
# 
# sync_gam_ca <- gam(Sync ~ s(cos_Angle), data = comp_data_near_3)
# sync_gam_ca_f <- gam(Sync ~ s(cos_Angle,by=Flow), data = comp_data_near_3)
# sync_gam_ca_a <- gam(Sync ~ s(cos_Angle,by=Ablation), data = comp_data_near_3)
# sync_gam_ca_d <- gam(Sync ~ s(cos_Angle,by=Darkness), data = comp_data_near_3)
# sync_gam_ca_fa <- gam(Sync ~ s(cos_Angle,by=Flow_Ablation), data = comp_data_near_3)
# sync_gam_ca_fd <- gam(Sync ~ s(cos_Angle,by=Flow_Darkness), data = comp_data_near_3)
# sync_gam_ca_fda <- gam(Sync ~ s(cos_Angle,by=Flow_Ablation_Darkness), data = comp_data_near_3)
# 
# sync_gam_sa <- gam(Sync ~ s(sin_Angle), data = comp_data_near_3)
# sync_gam_sa_f <- gam(Sync ~ s(sin_Angle,by=Flow), data = comp_data_near_3)
# sync_gam_sa_a <- gam(Sync ~ s(sin_Angle,by=Ablation), data = comp_data_near_3)
# sync_gam_sa_d <- gam(Sync ~ s(sin_Angle,by=Darkness), data = comp_data_near_3)
# sync_gam_sa_fa <- gam(Sync ~ s(sin_Angle,by=Flow_Ablation), data = comp_data_near_3)
# sync_gam_sa_fd <- gam(Sync ~ s(sin_Angle,by=Flow_Darkness), data = comp_data_near_3)
# sync_gam_sa_fda <- gam(Sync ~ s(sin_Angle,by=Flow_Ablation_Darkness), data = comp_data_near_3)
# 
# sync_gam_dca <- gam(Sync ~ s(Distance,cos_Angle), data = comp_data_near_3)
# sync_gam_dca_f <- gam(Sync ~ s(Distance,cos_Angle,by=Flow), data = comp_data_near_3)
# sync_gam_dca_a <- gam(Sync ~ s(Distance,cos_Angle,by=Ablation), data = comp_data_near_3)
# sync_gam_dca_d <- gam(Sync ~ s(Distance,cos_Angle,by=Darkness), data = comp_data_near_3)
# sync_gam_dca_fa <- gam(Sync ~ s(Distance,cos_Angle,by=Flow_Ablation), data = comp_data_near_3)
# sync_gam_dca_fd <- gam(Sync ~ s(Distance,cos_Angle,by=Flow_Darkness), data = comp_data_near_3)
# sync_gam_dca_fda <- gam(Sync ~ s(Distance,cos_Angle,by=Flow_Ablation_Darkness), data = comp_data_near_3)
# 
# sync_gam_dsa <- gam(Sync ~ s(Distance,sin_Angle), data = comp_data_near_3)
# sync_gam_dsa_f <- gam(Sync ~ s(Distance,sin_Angle,by=Flow), data = comp_data_near_3)
# sync_gam_dsa_a <- gam(Sync ~ s(Distance,sin_Angle,by=Ablation), data = comp_data_near_3)
# sync_gam_dsa_d <- gam(Sync ~ s(Distance,sin_Angle,by=Darkness), data = comp_data_near_3)
# sync_gam_dsa_fa <- gam(Sync ~ s(Distance,sin_Angle,by=Flow_Ablation), data = comp_data_near_3)
# sync_gam_dsa_fd <- gam(Sync ~ s(Distance,sin_Angle,by=Flow_Darkness), data = comp_data_near_3)
# sync_gam_dsa_fda <- gam(Sync ~ s(Distance,sin_Angle,by=Flow_Ablation_Darkness), data = comp_data_near_3)
# 
# sync_gam_dcasa <- gam(Sync ~ s(Distance,cos_Angle,sin_Angle), data = comp_data_near_3)
# sync_gam_dcasa_f <- gam(Sync ~ s(Distance,cos_Angle,sin_Angle,by=Flow), data = comp_data_near_3)
# sync_gam_dcasa_a <- gam(Sync ~ s(Distance,cos_Angle,sin_Angle,by=Ablation), data = comp_data_near_3)
# sync_gam_dcasa_d <- gam(Sync ~ s(Distance,cos_Angle,sin_Angle,by=Darkness), data = comp_data_near_3)
# sync_gam_dcasa_fa <- gam(Sync ~ s(Distance,cos_Angle,sin_Angle,by=Flow_Ablation), data = comp_data_near_3)
# sync_gam_dcasa_fd <- gam(Sync ~ s(Distance,cos_Angle,sin_Angle,by=Flow_Darkness), data = comp_data_near_3)
# sync_gam_dcasa_fda <- gam(Sync ~ s(Distance,cos_Angle,sin_Angle,by=Flow_Ablation_Darkness), data = comp_data_near_3)

Sync_AIC <- AIC(sync_gam_d,sync_gam_d_f,sync_gam_d_a,sync_gam_d_d,sync_gam_d_fa,sync_gam_d_fd,sync_gam_d_fda,
                 sync_gam_ca,sync_gam_ca_f,sync_gam_ca_a,sync_gam_ca_d,sync_gam_ca_fa,sync_gam_ca_fd,sync_gam_ca_fda,
                 sync_gam_sa,sync_gam_sa_f,sync_gam_sa_a,sync_gam_sa_d,sync_gam_sa_fa,sync_gam_sa_fd,sync_gam_sa_fda,
                 sync_gam_dca,sync_gam_dca_f,sync_gam_dca_a,sync_gam_dca_d,sync_gam_dca_fa,sync_gam_dca_fd,sync_gam_dca_fda,
                 sync_gam_dsa,sync_gam_dsa_f,sync_gam_dsa_a,sync_gam_dsa_d,sync_gam_dsa_fa,sync_gam_dsa_fd,sync_gam_dsa_fda,
                 sync_gam_dcasa,sync_gam_dcasa_f,sync_gam_dcasa_a,sync_gam_dcasa_d,sync_gam_dcasa_fa,sync_gam_dcasa_fd,sync_gam_dcasa_fda)

Sync_AIC <- tibble::rownames_to_column(Sync_AIC, "Model_Name")

Sync_AIC <- Sync_AIC %>% mutate(Model_Name = fct_reorder(Model_Name, -AIC)) %>% arrange(AIC)

ggplot(Sync_AIC %>% arrange(AIC), aes(x = Model_Name, y = AIC)) +
  geom_point(color = "blue") +
  xlab("") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme_light()+
  scale_y_reverse() +
  coord_flip()

Sync_AIC
```

Now we use the best of those
```{r}
speed_gam <- speed_gam_dsa_fda <- gam(Speed_Diff ~ s(Distance,sin_Angle,by=Flow_Ablation_Darkness), data = comp_data_near_3)
summary(speed_gam)
vis.gam(speed_gam)
# speed_gam_viz <- getViz(speed_gam)
# plotRGL(sm(speed_gam_viz, 1))

cos_heading_gam <- gam(cos_Heading ~ s(Distance,cos_Angle,sin_Angle,by=Flow_Ablation_Darkness), data = comp_data_near_3)
summary(cos_heading_gam)
vis.gam(cos_heading_gam)
# cos_heading_gam_viz <- getViz(cos_heading_gam)
# plotRGL(sm(cos_heading_gam_viz, 1), fix = c("Distance" = 0))

sin_heading_gam <- gam(sin_Heading ~ s(Distance,cos_Angle,sin_Angle,by=Flow_Ablation_Darkness), data = comp_data_near_3)
summary(sin_heading_gam)
vis.gam(sin_heading_gam)
# sin_heading_gam_viz <- getViz(sin_heading_gam)
# plotRGL(sm(sin_heading_gam_viz, 1), fix = c("Distance" = 0))

sync_gam <- gam(Sync ~ s(Distance,cos_Angle,sin_Angle,by=Flow_Ablation_Darkness), data = comp_data_near_3)
summary(sync_gam)
vis.gam(sync_gam)
# sync_gam_viz <- getViz(sync_gam)
# plotRGL(sm(sync_gam_viz, 1), fix = c("Distance" = 0))
```

Now we make some predictions for this.
```{r}
d <- seq(from = 0, to = 3, by = 0.1)
a <- seq(from = 0, to = 180, by = 5)

flows <- c("Flow 0", "Flow 2")
ablation <- c("No Ablation", "Ablated")
dark <- c("Light","Dark")

predict_df <- expand.grid(Distance = d, Angle = a, Flow = flows, Ablation = ablation, Darkness = dark)

predict_df <- predict_df %>% mutate(Flow_Ablation_Darkness = factor(paste(Flow,Ablation,Darkness,sep=", "))) %>% 
                             filter(!(Ablation == "Ablated" & Darkness == 'Dark')) %>%
                             mutate(rad_Angle = deg2rad(Angle)) %>%
                             mutate(sin_Angle = sin(rad_Angle), cos_Angle = cos(rad_Angle))

predict_df <- predict_df %>% mutate(pred_speed_diff = predict.gam(speed_gam,predict_df),
                                    pred_cos_heading_diff = predict.gam(cos_heading_gam,predict_df),
                                    pred_sin_heading_diff = predict.gam(sin_heading_gam,predict_df),
                                    pred_sync = predict.gam(sync_gam,predict_df)) %>%
                                    mutate(Angle = rad2deg(atan2(sin_Angle,cos_Angle))) %>%
                                    mutate(pred_heading_diff = rad2deg(atan2(pred_sin_heading_diff,pred_cos_heading_diff)))

predict_df_sum <- predict_df %>% group_by(Distance,Flow,Ablation,Darkness) %>%
                             mutate(mean_Speed_Distance = mean(pred_speed_diff), sd_Speed_Distance = sd(pred_speed_diff),
                                    mean_Heading_Distance = mean(pred_heading_diff), sd_Heading_Distance = sd(pred_heading_diff),
                                    mean_Sync_Distance = mean(pred_sync), sd_Sync_Distance = sd(pred_sync)) %>%
                             ungroup() %>%
                             group_by(Angle,Flow,Ablation,Darkness) %>%
                             mutate(mean_Speed_Angle = mean(pred_speed_diff), sd_Speed_Angle = sd(pred_speed_diff),
                                    mean_Heading_Angle = rad2deg(atan(mean(pred_sin_heading_diff)/mean(pred_cos_heading_diff))),
                                    sd_Heading_Angle = rad2deg(atan(sd(pred_sin_heading_diff)/sd(pred_cos_heading_diff))),
                                    mean_Sync_Angle = mean(pred_sync), sd_Sync_Angle = sd(pred_sync)) %>%
                             ungroup()
```

And now we graph those?

```{r}
ggplot(predict_df_sum, aes(x = Distance, y = mean_Speed_Distance, color = interaction(Ablation,Darkness,sep=", "),
                       fill = interaction(Ablation,Darkness,sep=" "))) +
  geom_line() +
  #geom_point(data=predict_df, aes(x = Distance, y = pred_speed_diff, alpha = 0.15)) +
  geom_ribbon(aes(ymin = mean_Speed_Distance-sd_Speed_Distance, ymax = mean_Speed_Distance+sd_Speed_Distance),
             linetype = 0,
             alpha = 0.1) +
  facet_wrap(~ Flow) +
  guides(fill = "none", color = guide_legend(title = "Condition")) +
  scale_color_manual(values=c("#000000", "#4CB944", "#00A7E1")) +
  scale_fill_manual(values=c("#000000", "#4CB944", "#00A7E1")) +
  ggtitle("Effect of Distance and Senses on Speed Difference") +
  xlab("Distance (BL)") +
  ylab("Speed Difference (BL/s)") +
  #ylim(0,1.5) +
  theme_light()

ggplot(predict_df_sum, aes(x = Distance, y = mean_Heading_Distance, color = interaction(Ablation,Darkness,sep=", "),
                       fill = interaction(Ablation,Darkness,sep=", "))) +
  geom_line() +
  #geom_point(data=predict_df, aes(x = Distance, y = pred_heading_diff, alpha = 0.15)) +
  geom_ribbon(aes(ymin = mean_Heading_Distance-sd_Heading_Distance, ymax = mean_Heading_Distance+sd_Heading_Distance),
              linetype = 0,
              alpha = 0.1) +
  facet_wrap(~ Flow) +
  guides(fill = "none", color = guide_legend(title = "Condition")) +
  scale_color_manual(values=c("#000000", "#4CB944", "#00A7E1")) +
  scale_fill_manual(values=c("#000000", "#4CB944", "#00A7E1")) +
  ggtitle("Effect of Distance and Senses on Heading Difference") +
  xlab("Distance (BL)") +
  ylab("Heading Difference (Degrees)") +
  #ylim(0,50) +
  theme_light()

ggplot(predict_df_sum, aes(x = Distance, y = mean_Sync_Distance, color = interaction(Ablation,Darkness,sep=", "),
                       fill = interaction(Ablation,Darkness,sep=", "))) +
  geom_line() +
  #geom_point(data=predict_df, aes(x = Distance, y = pred_sync, alpha = 0.15)) +
  geom_ribbon(aes(ymin = mean_Sync_Distance-sd_Sync_Distance, ymax = mean_Sync_Distance+sd_Sync_Distance),
              linetype = 0,
              alpha = 0.1) +
  facet_wrap(~ Flow) +
  guides(fill = "none", color = guide_legend(title = "Condition")) +
  scale_color_manual(values=c("#000000", "#4CB944", "#00A7E1")) +
  scale_fill_manual(values=c("#000000", "#4CB944", "#00A7E1")) +
  ggtitle("Effect of Distance and Senses on Synchonization") +
  xlab("Distance (BL)") +
  ylab("Synchonization") +
  #ylim(0,1)+
  theme_light()
```
```{r}
ggplot(predict_df_sum, aes(x = Angle, y = mean_Speed_Angle, color = interaction(Ablation,Darkness,sep=", "),
                       fill = interaction(Ablation,Darkness,sep=", "))) +
  geom_line() +
  #geom_point(data=predict_df, aes(x = Angle, y = pred_speed_diff, alpha = 0.15)) +
  geom_ribbon(aes(ymin = mean_Speed_Angle-sd_Speed_Angle, ymax = mean_Speed_Angle+sd_Speed_Angle),
              linetype = 0,
              alpha = 0.1) +
  facet_wrap(~ Flow) +
  guides(fill = "none", color = guide_legend(title = "Condition")) +
  scale_color_manual(values=c("#000000", "#4CB944", "#00A7E1")) +
  scale_fill_manual(values=c("#000000", "#4CB944", "#00A7E1")) +
  ggtitle("Effect of Bearing and Senses on Speed Difference") +
  xlab("Bearing (Degrees)") +
  ylab("Speed Difference (BL/s)") +
  theme_light()

ggplot(predict_df_sum, aes(x = Angle, y = mean_Heading_Angle, color = interaction(Ablation,Darkness,sep=", "),
                       fill = interaction(Ablation,Darkness,sep=", "))) +
  geom_line() +
  #geom_point(data=predict_df, aes(x = Angle, y = pred_heading_diff, alpha = 0.15)) +
  geom_ribbon(aes(ymin = mean_Heading_Angle-sd_Heading_Angle, ymax = mean_Heading_Angle+sd_Heading_Angle),
              linetype = 0,
              alpha = 0.1) +
  facet_wrap(~ Flow) +
  guides(fill = "none", color = guide_legend(title = "Condition")) +
  scale_color_manual(values=c("#000000", "#4CB944", "#00A7E1")) +
  scale_fill_manual(values=c("#000000", "#4CB944", "#00A7E1")) +
  ggtitle("Effect of Bearing and Senses on Heading Difference") +
  xlab("Angle (Degrees)") +
  ylab("Heading Difference (Degrees)") +
  theme_light()

ggplot(predict_df_sum, aes(x = Angle, y = mean_Sync_Angle, color = interaction(Ablation,Darkness,sep=", "),
                       fill = interaction(Ablation,Darkness,sep=", "))) +
  geom_line() +
 #geom_point(data=predict_df, aes(x = Angle, y = pred_sync, alpha = 0.15)) +
  geom_ribbon(aes(ymin = mean_Sync_Angle-sd_Sync_Angle, ymax = mean_Sync_Angle+sd_Sync_Angle),
              linetype = 0,
              alpha = 0.1) +
  facet_wrap(~ Flow) +
  guides(fill = "none", color = guide_legend(title = "Condition")) +
  scale_color_manual(values=c("#000000", "#4CB944", "#00A7E1")) +
  scale_fill_manual(values=c("#000000", "#4CB944", "#00A7E1")) +
  ggtitle("Effect of Bearing and Senses on Synchonization") +
  xlab("Bearing (Degrees)") +
  ylab("Synchonization") +
  theme_light()
```
Density vs Things

```{r}
dist_bin_size <- 1
angle_bin_size <- 30

comp_data_freqs <- comp_data_near_3 %>% mutate(Dist_bin = round_any(Distance, dist_bin_size), 
                                               Angle_bin = round_any(Angle, angle_bin_size)) %>%
                                        group_by(Flow,Darkness,Ablation, Dist_bin, Angle_bin,Flow_Ablation_Darkness) %>%
                                        mutate(Count = n()) %>%
                                        ungroup() %>%
                                        group_by(Flow,Darkness,Ablation,Flow_Ablation_Darkness) %>%
                                        mutate(Freq = Count/n()) %>%
                                        ungroup() %>%
                                        group_by(Flow,Darkness,Ablation,Dist_bin,Angle_bin,Flow_Ablation_Darkness) %>%
                                        summarise(Speed_Diff = mean(Speed_Diff),
                                                  Fold_Heading_Diff = ang_mean(fold_heading_diff),
                                                  Sync = mean(Sync),
                                                  Freq = mean(Freq))

```

```{r}
freq_speed_glm <- glm(Speed_Diff ~ Freq*(Flow+Darkness+Ablation+Flow:Ablation+Flow:Darkness), data = comp_data_freqs)

summary(freq_speed_glm)

freq_heading_glm <- glm(Fold_Heading_Diff ~ Freq*(Flow+Darkness+Ablation+Flow:Ablation+Flow:Darkness), data = comp_data_freqs)

summary(freq_heading_glm)

freq_sync_glm <- glm(Sync ~ Freq*(Flow+Darkness+Ablation+Flow:Ablation+Flow:Darkness), data = comp_data_freqs)

summary(freq_sync_glm)

freq_speed_gam <- gam(Speed_Diff ~ s(Freq,by=Flow_Ablation_Darkness),
                              data = comp_data_freqs)

summary(freq_speed_gam)

freq_heading_gam <- gam(Fold_Heading_Diff ~ s(Freq,by=Flow_Ablation_Darkness),
                              data = comp_data_freqs)

summary(freq_heading_gam)

freq_sync_gam <- gam(Sync ~ s(Freq,by=Flow_Ablation_Darkness),
                              data = comp_data_freqs)

summary(freq_sync_gam)
```

Now we make some predictions for this.
```{r}
f <- seq(from = 0.005, to = 0.07, by = 0.001)

flows <- c("Flow 0", "Flow 2")
ablation <- c("No Ablation", "Ablated")
dark <- c("Light","Dark")

freq_predict_df <- expand.grid(Freq = f, Flow = flows, Ablation = ablation, Darkness = dark)

freq_predict_df <- freq_predict_df %>% mutate(Flow_Ablation_Darkness = factor(paste(Flow,Ablation,Darkness,sep=", "))) %>% 
                            filter(!(Ablation == "Ablated" & Darkness == 'Dark'))

freq_predict_df <- freq_predict_df %>% mutate(pred_speed_diff = predict.gam(freq_speed_gam,freq_predict_df),
                                    pred_heading_diff = predict.gam(freq_heading_gam,freq_predict_df),
                                    pred_sync = predict.gam(freq_sync_gam,freq_predict_df))

```

```{r}
ggplot(comp_data_freqs, aes(x = Freq, y = Speed_Diff, color = interaction(Ablation,Darkness,sep=", "),
                       fill = interaction(Ablation,Darkness,sep=", "))) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
  facet_wrap(~ Flow) +
  guides(fill = "none", color = guide_legend(title = "Condition")) +
  scale_color_manual(values=c("#000000", "#4CB944", "#00A7E1")) +
  scale_fill_manual(values=c("#000000", "#4CB944", "#00A7E1")) +
  ggtitle("Effect of Preference and Senses on Speed Difference") +
  xlab("Frequency of Observations") +
  ylab("Speed Difference (BL/s)") +
  theme_light()

ggplot(comp_data_freqs, aes(x = Freq, y = Fold_Heading_Diff, color = interaction(Ablation,Darkness,sep=", "),
                       fill = interaction(Ablation,Darkness,sep=", "))) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
  facet_wrap(~ Flow) +
  guides(fill = "none", color = guide_legend(title = "Condition")) +
  scale_color_manual(values=c("#000000", "#4CB944", "#00A7E1")) +
  scale_fill_manual(values=c("#000000", "#4CB944", "#00A7E1")) +
  ggtitle("Effect of Preference and Senses on Heading Difference") +
  xlab("Frequency of Observations") +
  ylab("Heading Difference (Degrees)") +
  theme_light()

ggplot(comp_data_freqs, aes(x = Freq, y = Sync, color = interaction(Ablation,Darkness,sep=", "),
                       fill = interaction(Ablation,Darkness,sep=", "))) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", formula = "y ~ x", se = FALSE) +
  facet_wrap(~ Flow) +
  guides(fill = "none", color = guide_legend(title = "Condition")) +
  scale_color_manual(values=c("#000000", "#4CB944", "#00A7E1")) +
  scale_fill_manual(values=c("#000000", "#4CB944", "#00A7E1")) +
  ggtitle("Effect of Preference and Senses on Synchonization") +
  xlab("Frequency of Observations") +
  ylab("Synchonization") +
  theme_light()
```

```{r}
ggplot(freq_predict_df, aes(x = Freq, y = pred_speed_diff, color = interaction(Ablation,Darkness,sep=", "),
                       fill = interaction(Ablation,Darkness,sep=", "))) +
  geom_line() +
  facet_wrap(~ Flow) +
  guides(fill = "none", color = guide_legend(title = "Condition")) +
  scale_color_manual(values=c("#000000", "#4CB944", "#00A7E1")) +
  scale_fill_manual(values=c("#000000", "#4CB944", "#00A7E1")) +
  ggtitle("Effect of Frequency of Observations on Speed Difference") +
  xlab("Frequency (%)") +
  ylab("Speed Difference (BL/s)") +
  theme_light()

ggplot(freq_predict_df, aes(x = Freq, y = pred_heading_diff, color = interaction(Ablation,Darkness,sep=", "),
                       fill = interaction(Ablation,Darkness,sep=", "))) +
  geom_line() +
  facet_wrap(~ Flow) +
  guides(fill = "none", color = guide_legend(title = "Condition")) +
  scale_color_manual(values=c("#000000", "#4CB944", "#00A7E1")) +
  scale_fill_manual(values=c("#000000", "#4CB944", "#00A7E1")) +
  ggtitle("Effect of Frequency of Observations on Heading Difference") +
  xlab("Frequency (%)") +
  ylab("Heading Difference (Degrees)") +
  theme_light()

ggplot(freq_predict_df, aes(x = Freq, y = pred_sync, color = interaction(Ablation,Darkness,sep=", "),
                       fill = interaction(Ablation,Darkness,sep=", "))) +
  geom_line() +
  facet_wrap(~ Flow) +
  guides(fill = "none", color = guide_legend(title = "Condition")) +
  scale_color_manual(values=c("#000000", "#4CB944", "#00A7E1")) +
  scale_fill_manual(values=c("#000000", "#4CB944", "#00A7E1")) +
  ggtitle("Effect of Frequency of Observations on Synchonization") +
  xlab("Frequency (%)") +
  ylab("Synchonization") +
  theme_light()
```