### 2022-05-18 Fish Data Analysis/turning_fish_bias.Rmd

---
title: "Turning Fish Bias"
output: html_notebook
---

```{r}
library(tidyverse)
library(ggplot2)
library(bbmle) #For ICtab
library(car)
```

```{r}
turning_data <- read.csv("Annushka_Turning_Data_Wall.csv") %>%
                           filter(Next.to.Wall == 0) %>%
                           mutate(Flow = ifelse(Flow == "F0", "Still Water", "Flowing Water (2 BL/s)")) %>%
                           mutate(Ablation = ifelse(Ablation %in% c("LN","TN"), "No Ablation", "Ablated")) %>%
                           mutate(Darkness = ifelse(Darkness == "DN", "Light", "Dark")) %>%
                           mutate(Flow = factor(Flow), Ablation = factor(Ablation), Darkness = factor(Darkness)) %>%
                           mutate(Flow = fct_relevel(Flow, c("Still Water","Flowing Water (2 BL/s)"))) %>%
                           mutate(Ablation = fct_relevel(Ablation, c("No Ablation","Ablated"))) %>%
                           mutate(Darkness = fct_relevel(Darkness, c("Light","Dark"))) %>%
                           mutate(Flow_Ablation_Darkness = factor(paste(Flow,Ablation,Darkness,sep=", "))) %>%
                           mutate(Fish_Num_Ratio = No..Fish.Right/8) %>%
                           mutate(Fish_Ratio = paste(No..Fish.Left,No..Fish.Right,sep=":")) %>%
                           filter(Fish_Ratio %in% c("7:0","6:1","5:2","4:3","3:4","2:5","1:6","0:7")) %>%
                           mutate(Fish_Ratio = fct_relevel(Fish_Ratio,
                                                           c("7:0","6:1","5:2","4:3",
                                                             "3:4","2:5","1:6","0:7"))) %>%
                           mutate(turn_bool = ifelse(Turn.Direction == "right",1,0))

turning_data_sum <- turning_data %>% na.omit %>%
                                     group_by(Fish_Ratio,Fish_Num_Ratio,Flow,Ablation,Darkness,Flow_Ablation_Darkness) %>%
                                     summarize(mean_right_turn = mean(turn_bool),
                                               sd_right_turn = sd(turn_bool),
                                               se_right_turn = sd_right_turn/sqrt(n()),
                                               n = n()) %>%
                                     ungroup()

turning_data_sum_one <- turning_data %>% na.omit %>%
                                     group_by(Fish_Ratio,Fish_Num_Ratio) %>%
                                     summarize(mean_right_turn = mean(turn_bool),
                                               sd_right_turn = sd(turn_bool),
                                               se_right_turn = sd_right_turn/sqrt(n()),
                                               n = n()) %>%
                                     ungroup()

turning_data_count <- turning_data %>% na.omit %>%
                                     group_by(Fish_Ratio,Fish_Num_Ratio,Flow,Ablation,Darkness,Flow_Ablation_Darkness) %>%
                                     summarize(count_per = n()) %>%
                                     ungroup()
```

```{r}
ggplot(turning_data_sum, aes(x = Fish_Num_Ratio, y = mean_right_turn, color = Flow))+
  geom_point()+
  geom_line(alpha = 0.25) +
  # geom_errorbar(aes(ymin=mean_right_turn-se_right_turn,
  #                   ymax=mean_right_turn+se_right_turn),
  #               alpha = 0.25) +
  facet_wrap(~ Darkness + Ablation) +
  theme_classic() +
  xlab("# Fish on Left : # Fish on Right") +
  ylab("Probability of a Right Turn") +
  scale_colour_manual(values = c("Dark Grey","Blue")) +
  scale_x_continuous(breaks = c(0,0.125,0.25,0.375,
                                0.5,0.625,0.75,0.875),
                     labels = c("7:0","6:1","5:2","4:3",
                                "3:4","2:5","1:6","0:7"))

ggplot(turning_data_sum_one, aes(x = Fish_Num_Ratio, y = mean_right_turn))+
  geom_point()+
  geom_line(alpha = 0.25) +
  theme_classic() +
  xlab("# Fish on Left : # Fish on Right") +
  ylab("Probability of a Right Turn") +
  scale_x_continuous(breaks = c(0,0.125,0.25,0.375,
                                0.5,0.625,0.75,0.875),
                     labels = c("7:0","6:1","5:2","4:3",
                                "3:4","2:5","1:6","0:7"))

ggplot(turning_data_count, aes(x = Fish_Num_Ratio, y = count_per, fill = Flow))+
  geom_col(position = "dodge")+
  facet_wrap(~ Darkness + Ablation) +
  theme_classic() +
  xlab("# Fish on Left : # Fish on Right") +
  ylab("Count for Each Condition") +
  #scale_colour_manual(values = c("Dark Grey","Blue")) +
  scale_x_continuous(breaks = c(0,0.125,0.25,0.375,
                                0.5,0.625,0.75,0.875),
                     labels = c("7:0","6:1","5:2","4:3",
                                "3:4","2:5","1:6","0:7"))
```

```{r}
m0 <- glm(turn_bool ~ 1, data = turning_data, family = binomial)
m_fr <- glm(turn_bool ~ Fish_Num_Ratio, data = turning_data, family = binomial)
m_fl <- glm(turn_bool ~ Flow, data = turning_data, family = binomial)
m_a <- glm(turn_bool ~ Ablation, data = turning_data, family = binomial)
m_d <- glm(turn_bool ~ Darkness, data = turning_data, family = binomial)
ma_fr_fl <- glm(turn_bool ~ Fish_Num_Ratio + Flow, data = turning_data, family = binomial)
ma_fr_a <- glm(turn_bool ~ Fish_Num_Ratio + Ablation, data = turning_data, family = binomial)
ma_fr_d <- glm(turn_bool ~ Fish_Num_Ratio + Darkness, data = turning_data, family = binomial)
ma_fr_fl_a <- glm(turn_bool ~ Fish_Num_Ratio + Flow + Ablation, data = turning_data, family = binomial)
ma_fr_fl_d <- glm(turn_bool ~ Fish_Num_Ratio + Flow + Darkness, data = turning_data, family = binomial)
ma_fr_a_d <- glm(turn_bool ~ Fish_Num_Ratio + Ablation + Darkness, data = turning_data, family = binomial)
ma_fr_fl_a_d <- glm(turn_bool ~ Fish_Num_Ratio + Flow + Ablation + Darkness, data = turning_data, family = binomial)
mi_fr_fl <- glm(turn_bool ~ Fish_Num_Ratio * Flow, data = turning_data, family = binomial)
mi_fr_a <- glm(turn_bool ~ Fish_Num_Ratio * Ablation, data = turning_data, family = binomial)
mi_fr_d <- glm(turn_bool ~ Fish_Num_Ratio * Darkness, data = turning_data, family = binomial)
mi_fr_fl_a <- glm(turn_bool ~ Fish_Num_Ratio * Flow * Ablation, data = turning_data, family = binomial)
mi_fr_fl_d <- glm(turn_bool ~ Fish_Num_Ratio * Flow * Darkness, data = turning_data, family = binomial)
mi_fr_a_d <- glm(turn_bool ~ Fish_Num_Ratio * Ablation * Darkness, data = turning_data, family = binomial)
mi_fr_fl_a_d <- glm(turn_bool ~ Fish_Num_Ratio * Flow * Ablation * Darkness, data = turning_data, family = binomial)

ICtab(m0,m_fr,m_fl,m_a,m_d,
      ma_fr_fl,ma_fr_a,ma_fr_d,ma_fr_fl_a,ma_fr_fl_d,ma_fr_a_d,ma_fr_fl_a_d,
      mi_fr_fl,mi_fr_a,mi_fr_d,mi_fr_fl_a,mi_fr_fl_d,mi_fr_a_d,mi_fr_fl_a_d)

Anova(mi_fr_fl_a)

```

```{r}

#turning_data_graph <- turning_data %>% mutate(predicted = predict(mi_fr_d, newdata=turning_data, type="response"))

ggplot(turning_data, aes(x = Fish_Num_Ratio, y = turn_bool, color = Flow))+
  geom_jitter(alpha = 0.1, height = 0.025)+
  geom_line(aes(x = Fish_Num_Ratio, y = predict(ma_fr_fl_a, type="response"), color = Flow, shape = Ablation)) +
  geom_point(aes(x = Fish_Num_Ratio, y = predict(ma_fr_fl_a, type="response"), color = Flow, shape = Ablation)) +
  theme_classic() +
  xlab("# Fish on Left : # Fish on Right") +
  ylab("Probability of a Right Turn") +
  scale_colour_manual(values = c("#FF483E","#0C17A1")) +
  scale_x_continuous(breaks = c(0,0.125,0.25,0.375,
                                0.5,0.625,0.75,0.875),
                     labels = c("7:0","6:1","5:2","4:3",
                                "3:4","2:5","1:6","0:7"))

```

### NOW WITH SIMULATED DATA!

```{r}
turning_data_sim <- read.csv("Annushka_Turning_Data_Simmed.csv") %>% 
                           mutate(Flow = ifelse(Flow == "F0", "Still Water", "Flowing Water (2 BL/s)")) %>%
                           mutate(Ablation = ifelse(Ablation %in% c("LN","TN"), "No Ablation", ifelse(Ablation == "LY","Ablated",Ablation))) %>%
                           mutate(Darkness = ifelse(Darkness == "DN", "Light", ifelse(Darkness == "DY", "Dark", Darkness))) %>%
                           mutate(Flow = factor(Flow), Ablation = factor(Ablation), Darkness = factor(Darkness)) %>%
                           mutate(Flow = fct_relevel(Flow, c("Still Water","Flowing Water (2 BL/s)"))) %>%
                           mutate(Ablation = fct_relevel(Ablation, c("No Ablation","Ablated","Simulated"))) %>%
                           mutate(Darkness = fct_relevel(Darkness, c("Light","Dark","Simulated"))) %>%
                           mutate(Flow_Ablation_Darkness = factor(paste(Flow,Ablation,Darkness,sep=", "))) %>%
                           mutate(Fish_Num_Ratio = No..Fish.Right/8) %>%
                           mutate(Fish_Ratio = paste(No..Fish.Left,No..Fish.Right,sep=":")) %>%
                           filter(Fish_Ratio %in% c("7:0","6:1","5:2","4:3","3:4","2:5","1:6","0:7")) %>%
                           mutate(Fish_Ratio = fct_relevel(Fish_Ratio,
                                                           c("7:0","6:1","5:2","4:3",
                                                             "3:4","2:5","1:6","0:7"))) %>%
                           mutate(turn_bool = ifelse(Turn.Direction == "right",1,0))

turning_data_sim_sum <- turning_data_sim %>% na.omit %>%
                                     group_by(Fish_Ratio,Fish_Num_Ratio,Flow,Ablation,Darkness,Flow_Ablation_Darkness) %>%
                                     summarize(mean_right_turn = mean(turn_bool),
                                               sd_right_turn = sd(turn_bool),
                                               n = n()) %>%
                                     ungroup()

turning_data_sim_sum_one <- turning_data_sim %>% na.omit %>%
                                     group_by(Fish_Ratio,Fish_Num_Ratio) %>%
                                     summarize(mean_right_turn = mean(turn_bool),
                                               sd_right_turn = sd(turn_bool),
                                               n = n()) %>%
                                     ungroup()
```

```{r}
ggplot(turning_data_sim_sum, aes(x = Fish_Num_Ratio, y = mean_right_turn, color = Flow))+
  geom_point()+
  geom_line(alpha = 0.25) +
  # geom_errorbar(aes(ymin=mean_right_turn-sd_right_turn,
  #                   ymax=mean_right_turn+sd_right_turn),
  #               alpha = 0.25) +
  facet_wrap(~ Darkness + Ablation) +
  theme_classic() +
  xlab("# Fish on Left : # Fish on Right") +
  ylab("Probability of a Right Turn") +
  scale_colour_manual(values = c("Dark Grey","Blue")) +
  scale_x_continuous(breaks = c(0,0.125,0.25,0.375,
                                0.5,0.625,0.75,0.875),
                     labels = c("7:0","6:1","5:2","4:3",
                                "3:4","2:5","1:6","0:7"))

ggplot(turning_data_sum_one, aes(x = Fish_Num_Ratio, y = mean_right_turn))+
  geom_point()+
  geom_line(alpha = 0.25) +
  theme_classic() +
  xlab("# Fish on Left : # Fish on Right") +
  ylab("Probability of a Right Turn") +
  scale_x_continuous(breaks = c(0,0.125,0.25,0.375,
                                0.5,0.625,0.75,0.875),
                     labels = c("7:0","6:1","5:2","4:3",
                                "3:4","2:5","1:6","0:7"))
```

```{r}
m0 <- glm(turn_bool ~ 1, data = turning_data_sim, family = binomial)
m_fr <- glm(turn_bool ~ Fish_Num_Ratio, data = turning_data_sim, family = binomial)
m_fl <- glm(turn_bool ~ Flow, data = turning_data_sim, family = binomial)
m_a <- glm(turn_bool ~ Ablation, data = turning_data_sim, family = binomial)
m_d <- glm(turn_bool ~ Darkness, data = turning_data_sim, family = binomial)
ma_fr_fl <- glm(turn_bool ~ Fish_Num_Ratio + Flow, data = turning_data_sim, family = binomial)
ma_fr_a <- glm(turn_bool ~ Fish_Num_Ratio + Ablation, data = turning_data_sim, family = binomial)
ma_fr_d <- glm(turn_bool ~ Fish_Num_Ratio + Darkness, data = turning_data_sim, family = binomial)
ma_fr_fl_a <- glm(turn_bool ~ Fish_Num_Ratio + Flow + Ablation, data = turning_data_sim, family = binomial)
ma_fr_fl_d <- glm(turn_bool ~ Fish_Num_Ratio + Flow + Darkness, data = turning_data_sim, family = binomial)
ma_fr_a_d <- glm(turn_bool ~ Fish_Num_Ratio + Ablation + Darkness, data = turning_data_sim, family = binomial)
ma_fr_fl_a_d <- glm(turn_bool ~ Fish_Num_Ratio + Flow + Ablation + Darkness, data = turning_data_sim, family = binomial)
mi_fr_fl <- glm(turn_bool ~ Fish_Num_Ratio * Flow, data = turning_data_sim, family = binomial)
mi_fr_a <- glm(turn_bool ~ Fish_Num_Ratio * Ablation, data = turning_data_sim, family = binomial)
mi_fr_d <- glm(turn_bool ~ Fish_Num_Ratio * Darkness, data = turning_data_sim, family = binomial)
mi_fr_fl_a <- glm(turn_bool ~ Fish_Num_Ratio * Flow * Ablation, data = turning_data_sim, family = binomial)
mi_fr_fl_d <- glm(turn_bool ~ Fish_Num_Ratio * Flow * Darkness, data = turning_data_sim, family = binomial)
mi_fr_a_d <- glm(turn_bool ~ Fish_Num_Ratio * Ablation * Darkness, data = turning_data_sim, family = binomial)
mi_fr_fl_a_d <- glm(turn_bool ~ Fish_Num_Ratio * Flow * Ablation * Darkness, data = turning_data_sim, family = binomial)

ICtab(m0,m_fr,m_fl,m_a,m_d,
      ma_fr_fl,ma_fr_a,ma_fr_d,ma_fr_fl_a,ma_fr_fl_d,ma_fr_a_d,ma_fr_fl_a_d,
      mi_fr_fl,mi_fr_a,mi_fr_d,mi_fr_fl_a,mi_fr_fl_d,mi_fr_a_d,mi_fr_fl_a_d)

Anova(m_fr)
Anova(mi_fr_d)

```

```{r}

#turning_data_graph <- turning_data %>% mutate(predicted = predict(mi_fr_d, newdata=turning_data, type="response"))

ggplot(turning_data_sim, aes(x = Fish_Num_Ratio, y = turn_bool, color = Darkness))+
  geom_jitter(alpha = 0.1, height = 0.025)+
  geom_line(aes(x = Fish_Num_Ratio, y = predict(mi_fr_d, type="response"), color = Darkness)) +
  geom_point(aes(x = Fish_Num_Ratio, y = predict(mi_fr_d, type="response"), color = Darkness)) +
  theme_classic() +
  xlab("# Fish on Left : # Fish on Right") +
  ylab("Probability of a Right Turn") +
  scale_colour_manual(values = c("#FF483E","#0C17A1", "#777777")) +
  scale_x_continuous(breaks = c(0,0.125,0.25,0.375,
                                0.5,0.625,0.75,0.875),
                     labels = c("7:0","6:1","5:2","4:3",
                                "3:4","2:5","1:6","0:7"))

```