### 2023-02-09 Fish Data Analysis/Parallax Testing.qmd
Fixed code not pulling from correct V2 files
Code was only pulling from the base V2 file as I had a hard coded link in the middle. Now it uses the correct data files each time

---
title: "Parallax Testing"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(ggplot2)
library(bbmle) #For ICtab
library(car)
library(ggpubr)
library(ggdist)
library(emmeans)
library(viridis)

```

## Functions and Variables

```{r}
rad2deg <- function(rad) {(rad * 180) / (pi)}
deg2rad <- function(deg) {(deg * pi) / (180)}
round_any <- function(x, accuracy, f=round){f(x/ accuracy) * accuracy}
ang_mean <- function(x){rad2deg(atan2(mean(sin(deg2rad(x))),mean(cos(deg2rad(x)))))}

fold_angle_0_360_to_0_180 <- function(x){abs(abs(x-180)-180)}

fold_angle_neg_180_180_to_neg_90_90 <- function(x){ifelse(x > 90,90-abs(90-x),ifelse(x < -90, -90+abs(-90-x), x))}

min_n <- function(x,n){sort(x)[1:n]}
max_n <- function(x,n){sort(x,decreasing = TRUE)[1:n]}

light_no_ab_color = "#d4d7dd"
light_ab_color = "#f5ba9e"
dark_no_ab_color = "#5e94d4"
sim_color = "#777777"

flow_split_still = "#F59DE2"
flow_split_flow = "#5DD492"

dark_split_light = "#5e94d4"
dark_split_dark = "#011627"

tailbeat_len = 19

allowed_flow_type = c("Still Water","Flowing Water (2 BL/s)")

dot_binwidth = 0.0125
```

```{r}
data_2d <- read.csv("Data/Fish_Comp_Values.csv") %>% mutate(Flow = ifelse(Flow == "0", "Flow 0", "Flow 2")) %>%
                           mutate(Ablation = ifelse(Ablation == "N", "No Ablation", "Ablated")) %>%
                           mutate(Darkness = ifelse(Darkness == "N", "Light", "Dark")) %>%
                           mutate(Speed_Diff = abs(Speed_Diff)) %>%
                           mutate(Heading_Diff = abs(Heading_Diff)) %>%
                           mutate(quarter_heading_diff = abs(abs(Heading_Diff-90)-90)) %>%
                           mutate(Is_Aligned = ifelse(Heading_Diff < 30, 1, 0)) %>%
                           mutate(Is_Reversed = ifelse(Heading_Diff > 150, 1, 0))

data_3d_xy <- read.csv("Data/Fish_Comp_Values_3D_xy.csv") %>% mutate(Flow = ifelse(Flow == "0", "Flow 0", "Flow 2")) %>%
                           mutate(Ablation = ifelse(Ablation == "N", "No Ablation", "Ablated")) %>%
                           mutate(Darkness = ifelse(Darkness == "N", "Light", "Dark")) %>%
                           mutate(Speed_Diff = abs(Speed_Diff)) %>%
                           mutate(Heading_Diff = abs(Heading_Diff)) %>%
                           mutate(quarter_heading_diff = abs(abs(Heading_Diff-90)-90)) %>%
                           mutate(Is_Aligned = ifelse(Heading_Diff < 30, 1, 0)) %>%
                           mutate(Is_Reversed = ifelse(Heading_Diff > 150, 1, 0))
```

```{r}

data_2d_still_sum <- data_2d %>% filter(Darkness == "Light" & Ablation == "No Ablation") %>%
                                  mutate(round_fish1_x = round_any(Fish1_X/193,0.25),
                                         round_fish1_y = round_any(Fish1_Y/193,0.25)) %>%
                                  group_by(round_fish1_x,round_fish1_y) %>%
                                  summarise(mean_dist_2d = mean(Distance, na.rm = TRUE)) %>%
                                  ungroup()

data_3d_xy_still_sum <- data_3d_xy %>% filter(Darkness == "Light" & Ablation == "No Ablation") %>%
                                  mutate(round_fish1_x = round_any(Fish1_X/0.07862,0.25),
                                         round_fish1_y = round_any(Fish1_Y/0.07862,0.25)) %>%
                                  group_by(round_fish1_x,round_fish1_y) %>%
                                  summarise(mean_dist_3d_xy = mean(Distance, na.rm = TRUE)) %>%
                                  ungroup()

data_both <- full_join(data_3d_xy_still_sum, data_2d_still_sum) %>% na.omit() %>%
                      mutate(dist_diff = abs(mean_dist_3d_xy - mean_dist_2d))

```

Graphing NND in 2D
```{r}
ggplot(data_2d_still_sum, aes(round_fish1_x,round_fish1_y, color = mean_dist_2d))+
  geom_point(size = 4, shape = 15)+
  scale_color_viridis() +
  theme_classic() +
  coord_fixed()

ggplot(data_3d_xy_still_sum, aes(round_fish1_x,round_fish1_y, color = mean_dist_3d_xy))+
  geom_point(size = 2, shape = 15)+
  scale_color_viridis() +
  theme_classic() +
  coord_fixed()

ggplot(data_both, aes(round_fish1_x,round_fish1_y, color = dist_diff))+
  geom_point(size = 8, shape = 15)+
  scale_color_viridis() +
  theme_classic() +
  coord_fixed()

ggplot(data_both, aes(round_fish1_x,dist_diff))+
  geom_point()+
  geom_smooth(method = "glm", formula = "y ~ poly(x,2)")+ 
  theme_classic()

ggplot(data_both, aes(round_fish1_y,dist_diff))+
  geom_point()+
  geom_smooth(method = "glm", formula = "y ~ poly(x,2)")+
  theme_classic()
```

```{r}

xy_glm <- glm(dist_diff ~ round_fish1_x + round_fish1_y, data = data_both)
x2y_glm <- glm(dist_diff ~ I(round_fish1_x^2) + round_fish1_y, data = data_both)
xy2_glm <- glm(dist_diff ~ round_fish1_x + I(round_fish1_y^2), data = data_both)
xx2y_glm <- glm(dist_diff ~ round_fish1_x + I(round_fish1_x^2) + round_fish1_y, data = data_both)
xyy2_glm <- glm(dist_diff ~ round_fish1_x + round_fish1_y + I(round_fish1_y^2), data = data_both)
x2y2_glm <- glm(dist_diff ~ I(round_fish1_x^2) + I(round_fish1_y^2), data = data_both)

xx2y_m_glm <- glm(dist_diff ~ (round_fish1_x + I(round_fish1_x^2)) * round_fish1_y, data = data_both)

ICtab(xy_glm,x2y_glm,xy2_glm,xx2y_glm,xyy2_glm,x2y2_glm,xx2y_m_glm)

Anova(xx2y_m_glm)

data_both <- data_both %>% mutate(modeled = predict(xx2y_m_glm))

ggplot(data_both, aes(round_fish1_x,round_fish1_y, color = modeled))+
  geom_point(size = 8, shape = 15)+
  scale_color_viridis() +
  theme_classic() +
  coord_fixed()

```