### 2023-02-15 Charuco_DLT/rearrange_point_for_easywand.Rmd
Added Charuco Code

---
title: "R Notebook"
output: html_notebook
---

```{r setup}
library("tidyverse")
library("plotly")
library("imager")
library("here")
```

First step: Run `detect_charuco.ipynb` to detect the points in the videos of
the CHaRuCO boards.

Fill in the name of the output CSV file from `detect_charuco.ipynb`.
```{r}
detections <- read_csv("/Users/Ben/Desktop/Charuco_DLT/charuco_boards/2023_02_09/boards.csv") |> 
  select(-`...1`)
```

Fill in the directory that has the video files and the names of each video.
```{r}
videodir <- '/Users/Ben/Desktop/Charuco_DLT/charuco_boards/2023_02_09'
camfiles <- c('checkerboard_V1.mp4','checkerboard_V2.mp4')
camfiles <- map(camfiles, ~ file.path(videodir, .x))
```

NB: The code below uses `imager`, which won't load frames properly if there 
are any spaces in the path name anywhere

Make sure you set the video frame rate correctly.
```{r}
fps <- 20
```

This will show the 2D locations of each point for one camera.

```{r}
cam <- 2

detections |> 
  filter(camera == cam-1) |> 
  ggplot(aes(x = x, y = y, color = frame)) +
  geom_point()
```
This shows the range of x and y coordinates in the two cameras, just to check.
```{r}
detections |> 
  group_by(camera) |> 
  summarize(fr_hi = max(frame, na.rm = TRUE),
            across(c(x, y), 
                   list(lo = ~min(.x, na.rm = TRUE), hi = ~max(.x, na.rm = TRUE))))

```

This *should* load in a frame from one of the videos, but it doesn't always work, and I'm
not sure why.
```{r}
fr <- 10
cam <- 2

vidtime <- (fr-1) / fps

I1 <- imager::load.video(camfiles[[cam]], 
                         skip.to = vidtime,
                         frames = 1)
I1 <- as.data.frame(I1)
```

If the above block worked, then you should be able to plot the points and image
together here and check to make sure they look OK.
```{r}
detections |> 
  filter(frame == fr-1 & camera == cam-1) |> 
  ggplot(aes(x = x, y = y)) +
  geom_raster(data = I1, aes(x = x, y = y, fill = value)) +
  geom_point(aes(color = factor(id)))
```
# Set up points for EasyWand

To rearrange the points for EasyWand, we generate a fake "wand", which is any
pair of points that are separated by the same distance. This takes all of the pairs
that are separated by 1 grid square, which generates a *lot* of pairs. Not
every computer can process all of those points, so sometimes you want to generate
a smaller number. The "wand length" here is 1 grid square (=24.33 mm).
```{r}
pairs <- list(c(0,1), c(1,2), c(2,3), c(3,4),
              c(5,6), c(6,7), c(7,8), c(8,9),
              c(10,11), c(11,12), c(12,13), c(13,14),
              c(15,16), c(16,17), c(17,18), c(18,19),
              c(20,21), c(21,22), c(22,23), c(23,24),
              c(0,5), c(5,10), c(10,15), c(15,20),
              c(4,9), c(9,14), c(14,19), c(19,24))
```

This just gives the 4 pairs along the 4 edges of the grid. Then the "wand length" is
4 grid squares (=97.32mm).
```{r}
pairs <- list(c(0,4), c(4,24), c(24,20), c(20,0))
```

Generate all the pairs.
```{r}
mall <- list()
for (i in seq_along(pairs)) {
  p1 <- pairs[[i]]

  a1 <- detections |> filter(id == p1[1]) |> 
    rename(pt1_X = x, pt1_Y = y)
  a2 <- detections |> filter(id == p1[2]) |> 
    rename(pt2_X = x, pt2_Y = y)

  if (nrow(a1) > 0 & nrow(a2) > 0) {
    mall[[i]] = left_join(a1, a2, by = c('frame', 'camera')) |> 
      unite(id, c('id.x', 'id.y'))
  }
}

m <- bind_rows(mall)
```

Take a look at the beginning of the data frame, just to check.
```{r}
head(m)
```

Finish the reorganization.
```{r}
ptsforeasywand <-
  m |> 
  expand(camera, frame) |> 
  left_join(m, by = c('camera', 'frame')) |> 
  na.omit() |> 
  filter(!str_detect(id, 'NA')) |> 
  mutate(camera = camera+1) |> 
  pivot_wider(names_from = camera, values_from = starts_with('pt')) |> 
  rename_with(~str_replace(.x, '(pt\\d)_([X|Y])_(\\d)', '\\1_cam\\3_\\2')) |>
  select(pt1_cam1_X, pt1_cam1_Y,
         pt1_cam2_X, pt1_cam2_Y,
         pt2_cam1_X, pt2_cam1_Y,
         pt2_cam2_X, pt2_cam2_Y) |> 
  na.omit()

head(ptsforeasywand)
```

And save out the CSV.
```{r}
write_csv(ptsforeasywand, "/Users/Ben/Desktop/Charuco_DLT/2023_02_XX_calibrationcharuco_xypts.csv")
```

Final step: Load the CSV into EasyWand, punch in the correct fake "wand length",
add the camera parameters (resolution and focal length in pixels; check the SOPs
for 3D calibrations for these details for our cameras), and run the calibration.