### 2023-05-17 Fish Data Analysis/Raw Traces Figure.qmd
Fixing 2021_10_08_17 Data

---
title: "Raw Traces"
format: html
editor: visual
---

```{r}
library(gsignal)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(plotly)
library(zoo)
```

This is to make sure that it's the smae size after I do supsmu
```{r}
rad2deg <- function(rad) {(rad * 180) / (pi)}
deg2rad <- function(deg) {(deg * pi) / (180)}
round_any <- function(x, accuracy, f=round){f(x/ accuracy) * accuracy}
ang_mean <- function(x){rad2deg(atan2(mean(sin(deg2rad(x))),mean(cos(deg2rad(x)))))}

fold_angle_0_360_to_0_180 <- function(x){abs(abs(x-180)-180)}

fold_angle_neg_180_180_to_neg_90_90 <- function(x){ifelse(x > 90,90-abs(90-x),ifelse(x < -90, -90+abs(-90-x), x))}

supsmu_right_length_y <- function(x_data,y_data){
  output <- rep(NA,length(y_data))

  smooth_x <- supsmu(x_data,y_data)$x
  smooth_y <- supsmu(x_data,y_data)$y

  for(i in 1:length(smooth_x)){
    output[smooth_x[i]] <- smooth_y[i]
  }

  return(output)
}

fish_len <- 0.083197
```

Load the Data

```{r}
comp_data <- read.csv("Data/Fish_Raw_Comp_Values_3D.csv") %>% separate(Fish, c("Fish1", "Fish2"),sep="x") %>% mutate(Data_ID = paste(Year,Month,Day,Trial, sep = "_"))
```

```{r}
#comp_data <- comp_data %>% filter(Data_ID == "2021_8_3_17")

comp_data_switch_fish <- comp_data %>% mutate(Fish3 = Fish1,
                                              Fish3_X = Fish1_X,
                                              Fish3_Y = Fish1_Y,
                                              Fish3_Z = Fish1_Z,
                                              Fish3_Yaw_Heading = Fish1_Yaw_Heading,
                                              Fish3_Pitch_Heading = Fish1_Pitch_Heading,
                                              Fish3_Speed = Fish1_Speed) %>%
                                       mutate(Fish1 = Fish2,
                                              Fish1_X = Fish2_X,
                                              Fish1_Y = Fish2_Y,
                                              Fish1_Z = Fish2_Z,
                                              Fish1_Yaw_Heading = Fish2_Yaw_Heading,
                                              Fish1_Pitch_Heading = Fish2_Pitch_Heading,
                                              Fish1_Speed = Fish2_Speed) %>%
                                       mutate(Fish2 = Fish3,
                                              Fish2_X = Fish3_X,
                                              Fish2_Y = Fish3_Y,
                                              Fish2_Z = Fish3_Z,
                                              Fish2_Yaw_Heading = Fish3_Yaw_Heading,
                                              Fish2_Pitch_Heading = Fish3_Pitch_Heading,
                                              Fish2_Speed = Fish3_Speed) %>%
                                       select(-c(Fish3,Fish3_X,Fish3_Y,Fish3_Z,Fish3_Yaw_Heading,Fish3_Pitch_Heading,Fish3_Speed))

comp_data_two <- bind_rows(comp_data, comp_data_switch_fish)

comp_data_all <- comp_data_two %>% mutate(Flow = ifelse(Flow == "0", "Still Water", "Flowing Water (2 BL/s)")) %>%
                                   mutate(Ablation = ifelse(Ablation == "N", "No Ablation", "Ablated")) %>%
                                   mutate(Darkness = ifelse(Darkness == "N", "Light", "Dark")) %>%
                                   mutate(Flow = factor(Flow), Ablation = factor(Ablation), Darkness = factor(Darkness)) %>%
                                   mutate(Flow = fct_relevel(Flow, c("Still Water","Flowing Water (2 BL/s)"))) %>%
                                   mutate(Ablation = fct_relevel(Ablation, c("No Ablation","Ablated"))) %>%
                                   mutate(Darkness = fct_relevel(Darkness, c("Light","Dark"))) %>%
                                   mutate(Flow_Ablation_Darkness = factor(paste(Flow,Ablation,Darkness,sep=", "))) %>%
                                   mutate(Fish1_X = Fish1_X / fish_len,
                                          Fish1_Y = Fish1_Y / fish_len,
                                          Fish1_Z = Fish1_Z / fish_len,
                                          Fish2_X = Fish2_X / fish_len,
                                          Fish2_Y = Fish2_Y / fish_len,
                                          Fish2_Z = Fish2_Z / fish_len) #%>% 
                                   # mutate(Fish1_X = ifelse(Fish1_Speed > 10, NA, Fish1_X),
                                   #        Fish1_Y = ifelse(Fish1_Speed > 10, NA, Fish1_Y),
                                   #        Fish1_Speed = ifelse(Fish1_Speed > 10, NA, Fish1_Speed),
                                   #        Fish1_Yaw_Heading = ifelse(Fish1_Speed > 10, NA, Fish1_Yaw_Heading))

```

```{r}

unique((comp_data_all %>% filter(Darkness == "Dark"))$Data_ID)

unique(comp_data_all$Data_ID)

```

```{r}

"2020_7_28_3"
"2021_3_25_1"

# "2021_10_6_36" 4/5
# "2021_8_3_7" 2/5
# "2021_8_3_17" 4/5

# 2021_8_3_7:(206: 2 Groups)

#2021_7_7_27 ?

example_data <- comp_data_all %>% filter(Data_ID == "2021_10_8_27")

smooth_data <- example_data %>% group_by(Fish1,Frame_Num) %>%
                               summarise(Fish1_X = mean(Fish1_X),
                                         Fish1_Y = mean(Fish1_Y),
                                         Fish1_Z = mean(Fish1_Z),
                                         Fish1_Speed = mean(Fish1_Speed),
                                         Fish1_Yaw_Heading = mean(Fish1_Yaw_Heading)) %>%
                               group_by(Fish1) %>%
                               mutate(Fish1_X_Smooth = supsmu_right_length_y(Frame_Num,Fish1_X),
                                      Fish1_Y_Smooth = supsmu_right_length_y(Frame_Num,Fish1_Y),
                                      Fish1_Z_Smooth = supsmu_right_length_y(Frame_Num,Fish1_Z),
                                      Fish1_Speed_Smooth = supsmu_right_length_y(Frame_Num,Fish1_Speed),
                                      Speed_Increasing_Smooth = ifelse(Fish1_Speed < lead(Fish1_Speed),1,0),
                                      Fish1_Yaw_Heading_Max_Unwrap = rad2deg(supsmu_right_length_y(Frame_Num,unwrap(deg2rad(Fish1_Yaw_Heading))))) %>%
                               ungroup() %>% 
                               filter(Fish1_Z_Smooth >= -0.1 & Fish1_Z_Smooth <= 5)

smooth_data_nnd <- example_data %>% select(c(Fish1,Fish2,Frame_Num,Distance,Fish1_X,Fish1_Y,Fish2_X,Fish2_Y)) %>%
                                    group_by(Fish1,Frame_Num) %>%
                                    mutate(nndDistance = min(Distance, na.rm = TRUE),
                                           nndDistance = ifelse(is.infinite(nndDistance), NA, nndDistance),
                                           all_nas = all(is.na(nndDistance))) %>%
                                    filter(Distance == nndDistance | all_nas) %>%
                                    mutate(row_n = row_number()) %>%
                                    filter(row_n == 1) %>%
                                    ungroup() %>%
                                    group_by(Fish1) %>%
                                    arrange(Frame_Num) %>%
                                    mutate(Fish1_Distance_Smooth = supsmu_right_length_y(Frame_Num,nndDistance)) %>%
                                    ungroup()

nnd_summary <- smooth_data_nnd %>% group_by(Frame_Num) %>%
                                   summarise(meanNND = mean(nndDistance, na.rm = TRUE)) %>%
                                   ungroup() %>%
                                   mutate(meanNND_Smooth = supsmu_right_length_y(Frame_Num,meanNND))

nnd_joined <- example_data %>% select(c(Fish1,Fish2,Frame_Num,Distance,Fish1_X,Fish1_Y,Fish2_X,Fish2_Y)) %>%
                                    group_by(Fish1,Frame_Num) %>%
                                    mutate(nndDistance = min(Distance, na.rm = TRUE),
                                           nndDistance = ifelse(is.infinite(nndDistance), NA, nndDistance),
                                           all_nas = all(is.na(nndDistance))) %>%
                                    filter(Distance < 2 | all_nas) %>%
                                    mutate(row_n = row_number(),
                                           remove = ifelse(row_n > 1 & is.na(Distance),"Y","N")) %>%
                                    filter(remove == "N") %>%
                                    ungroup() %>%
                                    filter(Frame_Num %in% c(10)) %>% 
                                    mutate(Grouping = paste(Frame_Num, Fish1, Fish2)) %>% 
                                    select(c(Frame_Num, Fish1, Fish2, Grouping)) %>%
                                    pivot_longer(c(Fish1, Fish2),
                                                  names_to = "Useless",
                                                  values_to = "Fish1") %>%
                                    select(-c(Useless)) %>% 
                                    left_join(., smooth_data %>% 
                                    select(Fish1, Frame_Num, Fish1_X_Smooth, Fish1_Y_Smooth, Fish1_Z_Smooth))

polar_summary <- example_data %>% select(Fish1,Frame_Num,Fish1_Yaw_Heading) %>%
                                  group_by(Fish1,Frame_Num) %>%
                                  summarise(Fish1_Yaw_Heading = mean(Fish1_Yaw_Heading)) %>%
                                  ungroup() %>%
                                  mutate(sinHeading = sin(deg2rad(Fish1_Yaw_Heading)),
                                         cosHeading = cos(deg2rad(Fish1_Yaw_Heading))) %>%
                                  group_by(Frame_Num) %>%
                                  summarise(n = n(),
                                            sumSin = sum(sinHeading, na.rm = TRUE),
                                            sumCos =sum(cosHeading, na.rm = TRUE)) %>%
                                  ungroup() %>%
                                  mutate(polar = 1/n * sqrt(sumSin**2 + sumCos**2),
                                         polar_Smooth = supsmu_right_length_y(Frame_Num,polar))

raw_movement_plot <- ggplot()+
  geom_path(data = smooth_data, aes(x = Fish1_X_Smooth, y = Fish1_Y_Smooth, color = Fish1), size = 1)+
  geom_point(data = nnd_joined, aes(x = Fish1_X_Smooth, y = Fish1_Y_Smooth, color = Fish1), alpha = 1, size = 3)+
  geom_line(data = nnd_joined, aes(x = Fish1_X_Smooth, y = Fish1_Y_Smooth, fill = Grouping), alpha = 1)+
  theme_classic()+
  theme(legend.position = "none")

raw_speed_plot <- ggplot(smooth_data, aes(x = Frame_Num, y = Fish1_Speed_Smooth, color = Fish1))+
  geom_line(size = 1, alpha = 0.5)+
  theme_classic()+ 
  theme(legend.position = "none")

raw_heading_plot <- ggplot(smooth_data, aes(x = Frame_Num, y = Fish1_Yaw_Heading_Max_Unwrap, color = Fish1))+
  geom_line(size = 1, alpha = 0.5)+
  theme_classic()+ 
  theme(legend.position = "none")

raw_nnd_plot <- ggplot(smooth_data_nnd, aes(x = Frame_Num, y = Fish1_Distance_Smooth, color = Fish1))+
  geom_line(size = 1, alpha = 0.5)+
  geom_line(data = nnd_summary, aes(x = Frame_Num, y = meanNND_Smooth), color = "black", size = 1.5)+
  theme_classic()+ 
  theme(legend.position = "none")

raw_polar_plot <- ggplot(polar_summary, aes(x = Frame_Num, y = polar_Smooth))+
  geom_line(color = "black", size = 1.5)+
  theme_classic()+ 
  ylim(0,1)+
  theme(legend.position = "none")

raw_movement_plot
raw_speed_plot
raw_heading_plot
raw_nnd_plot
raw_polar_plot

raw_plots <- ggarrange(raw_movement_plot, raw_speed_plot, raw_heading_plot, raw_nnd_plot, raw_polar_plot,
          ncol = 1,
          labels = c("A","B","C","D","E"))

ggsave("Raw_Test.png", raw_plots, width = 10, height = 14)

fig <- plot_ly() %>%
       add_trace(x = smooth_data$Fish1_X_Smooth, y = smooth_data$Fish1_Y_Smooth, z = smooth_data$Fish1_Z_Smooth,
               split = smooth_data$Fish1, type = 'scatter3d', mode = 'lines', opacity = 1, line = list(width = 6),
               color = smooth_data$Fish1,
               colors = c("#F8766D","#CD9600","#7CAE00","#00BE67","#00BFC4","#00A9FF","#C77CFF","#FF61CC")) %>%

       add_trace(x = nnd_joined$Fish1_X_Smooth, y = nnd_joined$Fish1_Y_Smooth, z = nnd_joined$Fish1_Z_Smooth,
                 split = nnd_joined$Fish1, type = 'scatter3d', mode = "markers",
                 color = nnd_joined$Fish1,
                 colors = c("#F8766D","#CD9600","#7CAE00","#00BE67","#00BFC4","#00A9FF","#C77CFF","#FF61CC")) %>%

       add_trace(x = nnd_joined$Fish1_X_Smooth, y = nnd_joined$Fish1_Y_Smooth, z = nnd_joined$Fish1_Z_Smooth, 
                 split = nnd_joined$Grouping, type = 'scatter3d', mode = "lines", color = I("black")) %>%

       layout(showlegend = FALSE)

fig

```

 Okay now we're going to try and replace all missing parts with a spline

"2021_7_7_19"  "2021_7_7_27"  "2021_10_8_27" "2021_7_7_16"  "2021_7_7_17"  "2021_7_7_14"  "2021_10_8_21" "2021_10_8_26" "2021_7_7_6"   "2021_10_8_15"

```{r}

example_data_spline <- comp_data_all %>% filter(Data_ID == "2021_10_8_21")

spline_data <- example_data_spline %>% group_by(Fish1,Frame_Num) %>%
                               summarise(Fish1_X = mean(Fish1_X),
                                         Fish1_Y = mean(Fish1_Y),
                                         Fish1_Z = mean(Fish1_Z),
                                         Fish1_Speed = mean(Fish1_Speed),
                                         Fish1_Yaw_Heading = mean(Fish1_Yaw_Heading)) %>%
                               group_by(Fish1) %>%
                               mutate(Fish1_X_IB = ifelse(Fish1_Z >= -0.1 & Fish1_Z <= 5,Fish1_X,NA),
                                      Fish1_Y_IB = ifelse(Fish1_Z >= -0.1 & Fish1_Z <= 5,Fish1_Y,NA),
                                      Fish1_Z_IB = ifelse(Fish1_Z >= -0.1 & Fish1_Z <= 5,Fish1_Z,NA),
                                      Fish1_Speed_IB = ifelse(Fish1_Z >= -0.1 & Fish1_Z <= 5,Fish1_Speed,NA),
                                      Fish1_Heading_IB = ifelse(Fish1_Z >= -0.1 & Fish1_Z <= 5,Fish1_Yaw_Heading,NA)) %>%
                               mutate(Fish1_X_Spline = na.spline(Fish1_X_IB) + 0*na.approx(Fish1_X_IB, na.rm = FALSE),
                                      Fish1_Y_Spline = na.spline(Fish1_Y_IB) + 0*na.approx(Fish1_Y_IB, na.rm = FALSE),
                                      Fish1_Z_Spline = na.spline(Fish1_Z_IB) + 0*na.approx(Fish1_Z_IB, na.rm = FALSE),
                                      Fish1_Speed_Spline = na.spline(Fish1_Speed_IB) + 0*na.approx(Fish1_Speed_IB, na.rm = FALSE),
                                      Fish1_Heading_Spline = na.spline(Fish1_Heading_IB) + 0*na.approx(Fish1_Heading_IB, na.rm = FALSE))%>%
                               ungroup()

fig <- plot_ly() %>%
       add_trace(x = spline_data$Fish1_X_Spline, y = spline_data$Fish1_Y_Spline, z = spline_data$Fish1_Z_Spline,
               split = spline_data$Fish1, type = 'scatter3d', mode = 'lines', opacity = 1, line = list(width = 6),
               color = spline_data$Fish1,
               colors = c("#F8766D","#CD9600","#7CAE00","#00BE67","#00BFC4","#00A9FF","#C77CFF","#FF61CC")) %>%
       layout(showlegend = FALSE)

fig

```