### 2023-06-15 Fish Data Analysis/Heading Cross Corr.qmd
Working on heading Corrs

---
title: "Heading Cross Corr"
format: html
editor: visual
---

```{r}
library(gsignal)
library(tidyverse)
library(ggplot2)
library(bbmle) #For ICtab
library(car)
library(ggpubr)
library(ggdist)
library(emmeans)
library(lme4)
library(zoo)

library(conflicted)
conflict_prefer("filter", "dplyr")
```

```{r}
rad2deg <- function(rad) {(rad * 180) / (pi)}
deg2rad <- function(deg) {(deg * pi) / (180)}
round_any <- function(x, accuracy, f=round){f(x/ accuracy) * accuracy}
ang_mean <- function(x){rad2deg(atan2(mean(sin(deg2rad(x))),mean(cos(deg2rad(x)))))}

fold_angle_0_360_to_0_180 <- function(x){abs(abs(x-180)-180)}

fold_angle_neg_180_180_to_neg_90_90 <- function(x){ifelse(x > 90,90-abs(90-x),ifelse(x < -90, -90+abs(-90-x), x))}

min_n <- function(x,n){sort(x)[1:n]}
max_n <- function(x,n){sort(x,decreasing = TRUE)[1:n]}

light_no_ab_color = "#d4d7dd"
light_ab_color = "#f5ba9e"
dark_no_ab_color = "#5e94d4"
sim_color = "#777777"

flow_split_still = "#F59DE2"
flow_split_flow = "#5DD492"

dark_split_light = "#5e94d4"
dark_split_dark = "#011627"

tailbeat_len = 19

allowed_flow_type = c("Still Water","Flowing Water (2 BL/s)")

dot_binwidth = 0.03

text_size = 16
```

```{r}

x <- seq(0,2*pi,pi/10)
y1 <- sin(x)
y2 <- sin(x+pi)

plot(x,y1)
plot(x,y2)

ccf(y1, y2)

plot(ccf(1:10, 1:10)$lag,ccf(1:10, 1:10)$acf)

# plot(xcorr(1:10, 1:10, scale = "none")$lags,xcorr(1:10, 1:10, scale = "none")$R)
# plot(xcorr(1:10, 1:10, scale = "biased")$lags,xcorr(1:10, 1:10, scale = "biased")$R)
# plot(xcorr(1:10, 1:10, scale = "unbiased")$lags,xcorr(1:10, 1:10, scale = "unbiased")$R)
# plot(xcorr(1:10, 1:10, scale = "coeff")$lags,xcorr(1:10, 1:10, scale = "coeff")$R)

#let's try unbiased

```
Okay... let's try and get ccf to be flat wehn I'm coparing the same two things

Also here's the C code that calculates acf
Called from:

acf <- .Call(C_acf, x, lag.max, type == "correlation")

I think this translates it from R to C basically
so this is the call that goes to the function:

acf0(REAL(x), nx, ns, lagmax, cor, REAL(ans));

SEXP acf(SEXP x, SEXP lmax, SEXP sCor)
{
    int nx = nrows(x), ns = ncols(x), lagmax = asInteger(lmax),
	cor = asLogical(sCor);
    x = PROTECT(coerceVector(x, REALSXP));
    SEXP ans = PROTECT(allocVector(REALSXP, (lagmax + 1)*ns*ns));
    acf0(REAL(x), nx, ns, lagmax, cor, REAL(ans));
    SEXP d = PROTECT(allocVector(INTSXP, 3));
    INTEGER(d)[0] = lagmax + 1;
    INTEGER(d)[1] = INTEGER(d)[2] = ns;
    setAttrib(ans, R_DimSymbol, d);
    UNPROTECT(3);
    return ans;
}

x is the vector, n is the rows, ns is the columns, nl is the lag, cor is a boolean, acf is the output vector (called ans in SEXP acf)

static void
acf0(double *x, int n, int ns, int nl, Rboolean correlation, double *acf)
{
    int d1 = nl+1, d2 = ns*d1;

    for(int u = 0; u < ns; u++)
	for(int v = 0; v < ns; v++)
	    for(int lag = 0; lag <= nl; lag++) {
		double sum = 0.0; int nu = 0;
		for(int i = 0; i < n-lag; i++)
		    if(!ISNAN(x[i + lag + n*u]) && !ISNAN(x[i + n*v])) {
			nu++;
			sum += x[i + lag + n*u] * x[i + n*v];
		    }
		acf[lag + d1*u + d2*v] = (nu > 0) ? sum/(nu + lag) : NA_REAL;
	    }
    if(correlation) {
	if(n == 1) {
	    for(int u = 0; u < ns; u++)
		acf[0 + d1*u + d2*u] = 1.0;
	} else {
	    double *se = (double *) R_alloc(ns, sizeof(double));
	    for(int u = 0; u < ns; u++)
		se[u] = sqrt(acf[0 + d1*u + d2*u]);
	    for(int u = 0; u < ns; u++)
		for(int v = 0; v < ns; v++)
		    for(int lag = 0; lag <= nl; lag++) { // ensure correlations remain in  [-1,1] :
			double a = acf[lag + d1*u + d2*v] / (se[u]*se[v]);
			acf[lag + d1*u + d2*v] = (a > 1.) ? 1. : ((a < -1.) ? -1. : a);
		    }
	}
    }
}

I'm not 100% sure that the indexing and stuff is working exactly how I want it to so I'm going to 
just do it all via functions instead of trying to do mutate that cleanly

I think I only want lags starting at 0, no negative lags when we're matching up with distances

```{r}

get_corrs_lags <- function(f1_data, f2_data){

  output_len <- length(f1_data)

  output_vec <- rep(NA,output_len)

  saved_ccf <- ccf(f1_data, f2_data, na.action = na.pass, plot = F, lag.max = output_len)
  corrs <- saved_ccf$acf
  lags <- saved_ccf$lag

  pos_lags <- lags >= 0

  pos_lag_corrs <- corrs[pos_lags]
  pos_lag_lags <- lags[pos_lags]

  output_corrs <- c(pos_lag_corrs, output_vec)[1:output_len]
  output_lags <- c(pos_lag_lags, output_vec)[1:output_len]

  return(list(output_corrs,output_lags))
}

```

```{r}

fish_comp_data <- read.csv("Data/Fish_Comp_Values_3D.csv") %>% separate(Fish, c("Fish1", "Fish2"),sep="x") %>% 
                                                               mutate(Data_ID = paste(Year,Month,Day,Trial, sep = "_")) %>%
                                                               select(Year, Month, Day, Data_ID, Trial, Tailbeat_Num,
                                                                      Ablation, Darkness, Flow, Fish1, Fish2,
                                                                      Distance, Fish1_Yaw_Heading, Fish2_Yaw_Heading)

#fish_comp_data <- na.omit(fish_comp_data)

fish_comp_data_switch_fish <- fish_comp_data %>% mutate(Fish3 = Fish1,
                                                        Fish3_Yaw_Heading = Fish1_Yaw_Heading) %>%
                                       mutate(Fish1 = Fish2,
                                              Fish1_Yaw_Heading = Fish2_Yaw_Heading) %>%
                                       mutate(Fish2 = Fish3,
                                              Fish2_Yaw_Heading = Fish3_Yaw_Heading) %>%
                                       select(-c(Fish3,Fish3_Yaw_Heading))

fish_comp_data_combo <- bind_rows(fish_comp_data, fish_comp_data_switch_fish)

fish_comp_data_combo <- fish_comp_data_combo %>% mutate(Flow = ifelse(Flow == "0", "Still Water", "Flowing Water (2 BL/s)")) %>%
                           mutate(Ablation = ifelse(Ablation == "N", "No Ablation", "Ablated")) %>%
                           mutate(Darkness = ifelse(Darkness == "N", "Light", "Dark")) %>%
                           mutate(Flow = factor(Flow), Ablation = factor(Ablation), Darkness = factor(Darkness)) %>%
                           mutate(Flow = fct_relevel(Flow, c("Still Water","Flowing Water (2 BL/s)"))) %>%
                           mutate(Ablation = fct_relevel(Ablation, c("No Ablation","Ablated"))) %>%
                           mutate(Darkness = fct_relevel(Darkness, c("Light","Dark"))) %>%
                           mutate(Flow_Ablation_Darkness = factor(paste(Flow,Ablation,Darkness,sep="/n"))) %>%
                           arrange(Year,Month,Day,Trial,Fish1,Fish2,Tailbeat_Num)

fish_comp_data_combo <- fish_comp_data_combo %>% group_by(Year,Month,Day,Trial,Fish1,Fish2) %>%
                         mutate(Fish1_Yaw_Heading_Wrapped = Fish1_Yaw_Heading,
                                Fish2_Yaw_Heading_Wrapped = Fish2_Yaw_Heading,
                                Fish1_Yaw_Heading = unwrap(Fish1_Yaw_Heading),
                                Fish2_Yaw_Heading = unwrap(Fish2_Yaw_Heading))

```

Okay so we get the correlation for each time lag
And then have that to compare to the distances at that time
But how do you do a distance at a particular lag value? Do you just lag the coordinates and then calc the distances again?
That also doesn't seem right

```{r}

heading_corrs <- fish_comp_data_combo %>% group_by(Year,Month,Day,Trial,Fish1,Fish2) %>%
                                          mutate(len_run = sum(!is.na(Distance))) %>%
                                          ungroup() %>%
                                          filter(len_run > 2) %>%
                                          group_by(Year,Month,Day,Trial,Fish1,Fish2) %>%
                                          mutate(corrs = get_corrs_lags(Fish1_Yaw_Heading,Fish2_Yaw_Heading)[[1]],
                                                 lags = get_corrs_lags(Fish1_Yaw_Heading,Fish2_Yaw_Heading)[[2]])

heading_corrs_sum <- heading_corrs %>% filter(lags >= 0) %>%
                                       group_by(Year,Month,Day,Trial,Data_ID,Fish1,Fish2, Flow, Ablation, Darkness, Flow_Ablation_Darkness) %>%
                                       summarise(max_corr = max(corrs, na.rm = TRUE),
                                                 max_corr_lag = lags[na.omit(corrs == max_corr)],
                                                 lag_1_corr = corrs[lags == 1],
                                                 lag_1_distance = Distance[lags == 1],
                                                 max_corr_distance = Distance[na.omit(corrs == max_corr)]) %>%
                                       group_by(Flow, Ablation, Darkness) %>%
                                       mutate(Mean_Max_Corr = mean(max_corr, na.rm = T), SD_Max_Corr = sd(max_corr, na.rm = T),
                                              Mean_Max_Corr_Lag = mean(max_corr_lag, na.rm = T), SD_Max_Corr_Lag = sd(max_corr_lag, na.rm = T),
                                              Mean_Lag_1_Corr = mean(lag_1_corr, na.rm = T), SD_Lag_1_Corr = sd(lag_1_corr, na.rm = T),
                                              Mean_Lag_1_Distance = mean(lag_1_distance, na.rm = T), SD_Lag_1_Distance = sd(lag_1_distance, na.rm = T),
                                              Mean_Max_Corr_Distance = mean(max_corr_distance, na.rm = T), SD_Max_Corr_Distance = sd(max_corr_distance, na.rm = T)) %>%
                                       ungroup()
```

```{r}

ggplot(heading_corrs_sum %>% filter(Flow == "Still Water"),
       aes(x = interaction(Ablation,Darkness,sep=", "), y = max_corr,
           color = interaction(Ablation,Darkness,sep=", "),
           fill = interaction(Ablation,Darkness,sep=", ")))+
  stat_dots(side = "both", binwidth = unit(dot_binwidth/2, "npc")) +
  geom_point(aes(y = Mean_Max_Corr), size = 5, color = "Black") +
  geom_errorbar(aes(ymin=Mean_Max_Corr-SD_Max_Corr, ymax=Mean_Max_Corr+SD_Max_Corr), width=.2, color = "Black")+
  guides(fill = guide_legend(title = "Condition")) +
  scale_fill_manual(values=c(light_no_ab_color, light_ab_color, dark_no_ab_color)) +
  scale_color_manual(values=c(light_no_ab_color, light_ab_color, dark_no_ab_color)) +
  #ggtitle("Effect of Available Senses on Nearest Neighbor Distance (NND) in Still Water") +
  xlab("") +
  ylab("Max Correlation") +
  theme_light()+ 
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(size = text_size),
        axis.text.y = element_text(size = text_size),
        axis.title.y = element_text(size = text_size))+
  scale_x_discrete(labels=function(x){sub(",\\s", "\n", x)})

ggplot(heading_corrs_sum %>% filter(Flow == "Still Water"),
       aes(x = interaction(Ablation,Darkness,sep=", "), y = max_corr_lag,
           color = interaction(Ablation,Darkness,sep=", "),
           fill = interaction(Ablation,Darkness,sep=", ")))+
  stat_dots(side = "both", binwidth = unit(dot_binwidth/4, "npc")) +
  geom_point(aes(y = Mean_Max_Corr_Lag), size = 5, color = "Black") +
  geom_errorbar(aes(ymin=Mean_Max_Corr_Lag-SD_Max_Corr_Lag, ymax=Mean_Max_Corr_Lag+SD_Max_Corr_Lag), width=.2, color = "Black")+
  guides(fill = guide_legend(title = "Condition")) +
  scale_fill_manual(values=c(light_no_ab_color, light_ab_color, dark_no_ab_color)) +
  scale_color_manual(values=c(light_no_ab_color, light_ab_color, dark_no_ab_color)) +
  #ggtitle("Effect of Available Senses on Nearest Neighbor Distance (NND) in Still Water") +
  xlab("") +
  ylab("Lag at Max Correlation") +
  theme_light()+ 
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(size = text_size),
        axis.text.y = element_text(size = text_size),
        axis.title.y = element_text(size = text_size))+
  scale_x_discrete(labels=function(x){sub(",\\s", "\n", x)})

ggplot(heading_corrs_sum %>% filter(Flow == "Still Water"),
       aes(x = interaction(Ablation,Darkness,sep=", "), y = lag_1_corr,
           color = interaction(Ablation,Darkness,sep=", "),
           fill = interaction(Ablation,Darkness,sep=", ")))+
  stat_dots(side = "both", binwidth = unit(dot_binwidth/4, "npc")) +
  geom_point(aes(y = Mean_Lag_1_Corr), size = 5, color = "Black") +
  geom_errorbar(aes(ymin=Mean_Lag_1_Corr-SD_Lag_1_Corr, ymax=Mean_Lag_1_Corr+SD_Lag_1_Corr), width=.2, color = "Black")+
  guides(fill = guide_legend(title = "Condition")) +
  scale_fill_manual(values=c(light_no_ab_color, light_ab_color, dark_no_ab_color)) +
  scale_color_manual(values=c(light_no_ab_color, light_ab_color, dark_no_ab_color)) +
  #ggtitle("Effect of Available Senses on Nearest Neighbor Distance (NND) in Still Water") +
  xlab("") +
  ylab("Correlation at 1 Lag") +
  theme_light()+ 
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(size = text_size),
        axis.text.y = element_text(size = text_size),
        axis.title.y = element_text(size = text_size))+
  scale_x_discrete(labels=function(x){sub(",\\s", "\n", x)})

ggplot(heading_corrs_sum %>% filter(Flow == "Still Water"),
       aes(x = max_corr, y = max_corr_distance,
           color = interaction(Ablation,Darkness,sep=", "),
           fill = interaction(Ablation,Darkness,sep=", ")))+
  geom_point() +
  geom_smooth(se = F, method="glm") +
  guides(fill = guide_legend(title = "Condition")) +
  scale_color_manual(values=c(light_no_ab_color, light_ab_color, dark_no_ab_color)) +
  xlab("Max Correlation") +
  ylab("Distance at Max Correlation") +
  xlim(0,1)+
  theme_light()+ 
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(size = text_size),
        axis.text.y = element_text(size = text_size),
        axis.title.x = element_text(size = text_size),
        axis.title.y = element_text(size = text_size))

max_corr_vs_dist_glmer <- lmer(max_corr_distance ~ max_corr*(Ablation + Darkness) + (1|Data_ID), data = heading_corrs_sum %>% filter(Flow == "Still Water"))
Anova(max_corr_vs_dist_glmer)

```

# All the raw data

```{r}

raw_fish_comp_data <- read.csv("Data/Fish_Raw_Comp_Values_3D.csv") %>% separate(Fish, c("Fish1", "Fish2"),sep="x") %>% 
                                                               mutate(Data_ID = paste(Year,Month,Day,Trial, sep = "_")) %>%
                                                               select(Year, Month, Day, Data_ID, Trial, Frame_Num,
                                                                      Ablation, Darkness, Flow, Fish1, Fish2,
                                                                      Distance, Fish1_Yaw_Heading, Fish2_Yaw_Heading)

#fish_comp_data <- na.omit(fish_comp_data)

raw_fish_comp_data_switch_fish <- raw_fish_comp_data %>% mutate(Fish3 = Fish1,
                                                        Fish3_Yaw_Heading = Fish1_Yaw_Heading) %>%
                                       mutate(Fish1 = Fish2,
                                              Fish1_Yaw_Heading = Fish2_Yaw_Heading) %>%
                                       mutate(Fish2 = Fish3,
                                              Fish2_Yaw_Heading = Fish3_Yaw_Heading) %>%
                                       select(-c(Fish3,Fish3_Yaw_Heading))

raw_fish_comp_data_combo <- bind_rows(raw_fish_comp_data, raw_fish_comp_data_switch_fish)

raw_fish_comp_data_combo <- raw_fish_comp_data_combo %>% mutate(Flow = ifelse(Flow == "0", "Still Water", "Flowing Water (2 BL/s)")) %>%
                           mutate(Ablation = ifelse(Ablation == "N", "No Ablation", "Ablated")) %>%
                           mutate(Darkness = ifelse(Darkness == "N", "Light", "Dark")) %>%
                           mutate(Flow = factor(Flow), Ablation = factor(Ablation), Darkness = factor(Darkness)) %>%
                           mutate(Flow = fct_relevel(Flow, c("Still Water","Flowing Water (2 BL/s)"))) %>%
                           mutate(Ablation = fct_relevel(Ablation, c("No Ablation","Ablated"))) %>%
                           mutate(Darkness = fct_relevel(Darkness, c("Light","Dark"))) %>%
                           mutate(Flow_Ablation_Darkness = factor(paste(Flow,Ablation,Darkness,sep="/n"))) %>%
                           arrange(Year,Month,Day,Trial,Fish1,Fish2,Frame_Num)

raw_fish_comp_data_combo <- raw_fish_comp_data_combo %>% group_by(Year,Month,Day,Trial,Fish1,Fish2) %>%
                         mutate(Fish1_Yaw_Heading_Wrapped = Fish1_Yaw_Heading,
                                Fish2_Yaw_Heading_Wrapped = Fish2_Yaw_Heading,
                                Fish1_Yaw_Heading = unwrap(Fish1_Yaw_Heading),
                                Fish2_Yaw_Heading = unwrap(Fish2_Yaw_Heading))

```

Testing things out, I think I just needed a much bigger lag

```{r}

data_subset <- raw_fish_comp_data_combo %>% filter(Data_ID == "2021_10_6_36", Fish1 == "individual5", Fish2 == "individual2")

ggplot(data_subset, aes())+
  geom_point(aes(Frame_Num, Fish1_Yaw_Heading), color = "red")+
  geom_point(aes(Frame_Num, Fish2_Yaw_Heading), color = "blue")+
  theme_light()

ccf(data_subset$Fish1_Yaw_Heading, data_subset$Fish2_Yaw_Heading, na.action = na.pass, lag.max = length(data_subset$Fish1_Yaw_Heading))

```

So now we use a sliding window here instead of jsut the regular corr

Now make it move by 10 frames instead of 1 for the window

```{r}

#I want this function to return the highest correaltion and the lag for that correlation for each 
# window.

windowed_correlation_finder <- function(x,y){
  lag_max <- tailbeat_len
  window_size <- tailbeat_len*3

  window_step <- as.integer(tailbeat_len/2)

  window_starts <- rep(NA,(length(x)-window_size-lag_max)*lag_max)
  lags <- rep(NA,(length(x)-window_size-lag_max)*lag_max)
  corr <- rep(NA,(length(x)-window_size-lag_max)*lag_max)

  for (i in seq(1,length(x)-window_size-lag_max,window_step)) {

    for (j in 1:lag_max) {

      out_ind <- window_step*(i-1)+j

      fish1_window <- x[i:(i+window_size)]
      fish2_window <- y[(i+j):(i+j+window_size)]

      if(sum(!is.na(fish1_window+fish2_window)) > window_step){
        corr_val <- cor(fish1_window, fish2_window, method = "pearson", use = "complete.obs")

        window_starts[out_ind] <- i
        lags[out_ind] <- j
        corr[out_ind] <- corr_val
      }
      else{
        window_starts[out_ind] <- i
        lags[out_ind] <- j
        corr[out_ind] <- NA
      }

    }

  }

  all_data_frame <- tibble(window_start = window_starts,
                           lag = lags,
                           corr = corr)

  out_df <- all_data_frame %>% group_by(window_start) %>% 
                               mutate(isMax = (corr == max(corr, na.rm = TRUE))) %>% 
                               filter(isMax) %>% 
                               summarise(max_corr = first(corr), max_corr_lag = first(lag)) %>% 
                               ungroup()

  out_corrs <- out_df$max_corr
  out_lags <- out_df$max_corr_lag

  length(out_corrs) <- length(x)
  length(out_lags) <- length(x)

  output <- tibble(ccf = out_corrs, lags = out_lags)

  return(output)
}

```

Testing optimization
```{r}

x <- seq(0,2*pi,pi/10)
y1 <- sin(x)
y2 <- sin(x+pi/2 + rnorm(20,0,pi/4))

plot(x,y1)
plot(x,y2)

tester_window_size <- 5

y1_embed <- embed(y1,tester_window_size)
y2_embed <- embed(y2,tester_window_size)

list_embed <- data.frame(y1 = y1,y2 = y2)

cor_vals <- rollapply(list_embed, width = tester_window_size,
                      FUN = function(w) cor(list_embed[,1], list_embed[,2], method = "pearson", use = "complete.obs"),
                      align = "right")

cor_vals

z2 <- zoo(rnorm(6))
z2
rollapply(1:6, 3, mean)

```

Okay so now a specific correaltion version

```{r}
get_corrs_lags_ben <- function(f1_data, f2_data){

  output_len <- length(f1_data)

  output_vec <- rep(NA,output_len)

  saved_ccf <- windowed_correlation_finder(f1_data,f2_data)

  corrs <- saved_ccf$ccf
  lags <- saved_ccf$lags

  pos_lags <- lags >= 0

  pos_lag_corrs <- corrs[pos_lags]
  pos_lag_lags <- lags[pos_lags]

  output_corrs <- c(pos_lag_corrs, output_vec)[1:output_len]
  output_lags <- c(pos_lag_lags, output_vec)[1:output_len]

  return(list(output_corrs,output_lags))
}
```

Okay so we get the correlation for each time lag
And then have that to compare to the distances at that time
But how do you do a distance at a particular lag value? Do you just lag the coordinates and then calc the distances again?
That also doesn't seem right

```{r}

raw_heading_corrs <- raw_fish_comp_data_combo %>% #filter(Data_ID == "2021_8_3_11" & Fish1 == "individual8"  & Fish2 == "individual7") %>%
                                          group_by(Year,Month,Day,Trial,Fish1,Fish2) %>%
                                          mutate(len_run = sum(!is.na(Distance))) %>%
                                          ungroup() %>%
                                          filter(len_run > 2) %>%
                                          group_by(Year,Month,Day,Trial,Fish1,Fish2) %>%
                                          mutate(corrs = get_corrs_lags_ben(Fish1_Yaw_Heading,Fish2_Yaw_Heading)[[1]],
                                                 lags = get_corrs_lags_ben(Fish1_Yaw_Heading,Fish2_Yaw_Heading)[[2]])

raw_heading_corrs_sum <- raw_heading_corrs %>% filter(lags >= 0) %>%
                                       group_by(Year,Month,Day,Trial,Data_ID,Fish1,Fish2, Flow, Ablation, Darkness, Flow_Ablation_Darkness) %>%
                                       summarise(max_corr = first(max(corrs, na.rm = TRUE)),
                                                 max_corr_lag = first(lags[na.omit(corrs == max_corr)]),
                                                 max_corr_distance = first(Distance[na.omit(corrs == max_corr)])) %>%
                                       group_by(Flow, Ablation, Darkness) %>%
                                       mutate(Mean_Max_Corr = mean(max_corr, na.rm = T), SD_Max_Corr = sd(max_corr, na.rm = T),
                                              Mean_Max_Corr_Lag = mean(max_corr_lag, na.rm = T), SD_Max_Corr_Lag = sd(max_corr_lag, na.rm = T),
                                              Mean_Max_Corr_Distance = mean(max_corr_distance, na.rm = T), SD_Max_Corr_Distance = sd(max_corr_distance, na.rm = T)) %>%
                                       ungroup()
```

Okay so don't do a max of the time series, just plot all of the points, no mean and no max, just all distances and correlations

#graphing raw data

```{r}

ggplot(raw_heading_corrs_sum %>% filter(Flow == "Still Water"),
       aes(x = interaction(Ablation,Darkness,sep=", "), y = max_corr,
           color = interaction(Ablation,Darkness,sep=", "),
           fill = interaction(Ablation,Darkness,sep=", ")))+
  stat_dots(side = "both", binwidth = unit(dot_binwidth/2, "npc")) +
  geom_point(aes(y = Mean_Max_Corr), size = 5, color = "Black") +
  geom_errorbar(aes(ymin=Mean_Max_Corr-SD_Max_Corr, ymax=Mean_Max_Corr+SD_Max_Corr), width=.2, color = "Black")+
  guides(fill = guide_legend(title = "Condition")) +
  scale_fill_manual(values=c(light_no_ab_color, light_ab_color, dark_no_ab_color)) +
  scale_color_manual(values=c(light_no_ab_color, light_ab_color, dark_no_ab_color)) +
  ggtitle("Effect of Available Senses on Nearest Neighbor Distance (NND) in Still Water") +
  xlab("") +
  ylab("Max Correlation") +
  theme_light()+ 
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(size = text_size),
        axis.text.y = element_text(size = text_size),
        axis.title.y = element_text(size = text_size))+
  scale_x_discrete(labels=function(x){sub(",\\s", "\n", x)})

ggplot(raw_heading_corrs_sum %>% filter(Flow == "Still Water"),
       aes(x = interaction(Ablation,Darkness,sep=", "), y = max_corr_lag,
           color = interaction(Ablation,Darkness,sep=", "),
           fill = interaction(Ablation,Darkness,sep=", ")))+
  stat_dots(side = "both", binwidth = unit(dot_binwidth/4, "npc")) +
  geom_point(aes(y = Mean_Max_Corr_Lag), size = 5, color = "Black") +
  geom_errorbar(aes(ymin=Mean_Max_Corr_Lag-SD_Max_Corr_Lag, ymax=Mean_Max_Corr_Lag+SD_Max_Corr_Lag), width=.2, color = "Black")+
  guides(fill = guide_legend(title = "Condition")) +
  scale_fill_manual(values=c(light_no_ab_color, light_ab_color, dark_no_ab_color)) +
  scale_color_manual(values=c(light_no_ab_color, light_ab_color, dark_no_ab_color)) +
  #ggtitle("Effect of Available Senses on Nearest Neighbor Distance (NND) in Still Water") +
  xlab("") +
  ylab("Lag at Max Correlation") +
  theme_light()+ 
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(size = text_size),
        axis.text.y = element_text(size = text_size),
        axis.title.y = element_text(size = text_size))+
  scale_x_discrete(labels=function(x){sub(",\\s", "\n", x)})

ggplot(raw_heading_corrs_sum %>% filter(Flow == "Still Water"),
       aes(x = max_corr, y = max_corr_distance,
           color = interaction(Ablation,Darkness,sep=", "),
           fill = interaction(Ablation,Darkness,sep=", ")))+
  geom_point() +
  geom_smooth(se = F, method="glm") +
  guides(fill = guide_legend(title = "Condition")) +
  scale_color_manual(values=c(light_no_ab_color, light_ab_color, dark_no_ab_color)) +
  xlab("Max Correlation") +
  ylab("Distance at Max Correlation") +
  xlim(0,1)+
  theme_light()+ 
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(size = text_size),
        axis.text.y = element_text(size = text_size),
        axis.title.x = element_text(size = text_size),
        axis.title.y = element_text(size = text_size))

max_corr_glmer <- lmer(max_corr ~ Ablation + Darkness + (1|Data_ID), data = raw_heading_corrs_sum %>% filter(Flow == "Still Water"))
Anova(max_corr_glmer)

max_corr_vs_dist_glmer <- lmer(max_corr_distance ~ max_corr*(Ablation + Darkness) + (1|Data_ID), data = raw_heading_corrs_sum %>% filter(Flow == "Still Water"))
Anova(max_corr_vs_dist_glmer)

```