### 2024-03-15 Fish Data Analysis/NN_Accel_Heading.qmd
Single fish plotting works

---
title: "NN_Accel_Heading"
format: html
editor: visual
---

## Loading in the Libraries for Everything

```{r}
library(MASS)
library(tidyverse)
library(ggplot2)
library(bbmle) #For ICtab
library(car)
library(ggpubr)
library(ggdist)
library(emmeans)
library(lme4)
library(viridis)
library(lme4)

```

```{r}
rad2deg <- function(rad) {(rad * 180) / (pi)}
deg2rad <- function(deg) {(deg * pi) / (180)}
round_any <- function(x, accuracy, f=round){f(x/ accuracy) * accuracy}

deg_ang_mean <- function(x){rad2deg(atan2(mean(sin(deg2rad(x))),mean(cos(deg2rad(x)))))}
rad_ang_mean <- function(x){atan2(mean(sin(x)),mean(cos(x)))}

fold_angle_0_360_to_0_180 <- function(x){abs(abs(x-180)-180)}

fold_angle_neg_180_180_to_neg_90_90 <- function(x){ifelse(x > 90,90-abs(90-x),ifelse(x < -90, -90+abs(-90-x), x))}

min_n <- function(x,n){sort(x)[1:n]}
max_n <- function(x,n){sort(x,decreasing = TRUE)[1:n]}

nth_min <- function(x,n){sort(na.omit(x))[n]}

light_no_ab_color = "#d4d7dd"
light_ab_color = "#f5ba9e"
dark_no_ab_color = "#5e94d4"
dark_ab_color = "#1E4371"
sim_color = "#777777"

flow_split_still = "#F59DE2"
flow_split_flow = "#5DD492"

dark_split_light = "#5e94d4"
dark_split_dark = "#011627"

tailbeat_len = 19

allowed_flow_type = c("Still Water","Flowing Water (2 BL/s)")

dot_binwidth = 0.03

text_size = 16
```

Load in the data

```{r}

single_comp_data <- read.csv("Data/Fish_Comp_Values_3D_Single.csv") %>%
                           mutate(Flow = ifelse(Flow == "0", "Still Water", "Flowing Water (2 BL/s)")) %>%
                           mutate(Ablation = ifelse(Ablation  == "S", "Single", "Jingle")) %>%
                           mutate(Darkness = ifelse(Darkness == "S", "Single", "Jingle")) %>%
                           mutate(Flow = factor(Flow), Ablation = factor(Ablation), Darkness = factor(Darkness)) %>%
                           mutate(Flow = fct_relevel(Flow, c("Still Water"))) %>%
                           mutate(Ablation = fct_relevel(Ablation, c("Single"))) %>%
                           mutate(Darkness = fct_relevel(Darkness, c("Single"))) %>%
                           mutate(Flow_Ablation_Darkness = factor(paste(Flow,Ablation,Darkness,sep=", "))) %>%
                           separate(Fish, c("Fish1", "Fish2"),sep="x") %>% 
                           # # ### Used to remove every third point to reduce autocorrelation
                           mutate(Data_ID = paste(Year,Month,Day,Trial, sep = "_"),
                                  Date = paste(Year,Month,Day,Trial, sep = "_"),
                                  Fish_ID = paste(Year,Month,Day,Trial,Fish1, sep = "_"),
                                  Comp_ID = paste(Year,Month,Day,Trial,Fish1,Fish2, sep = "_")) %>%
                           group_by(Comp_ID) %>%
                           #filter(row_number() %% 3 == 1) %>%
                           ungroup() %>%
                           mutate(Type = "Single")

single_comp_data <- na.omit(single_comp_data)

fish_comp_data <- read.csv("Data/Fish_Comp_Values_3D.csv") %>%
                           mutate(Flow = ifelse(Flow == "0", "Still Water", "Flowing Water (2 BL/s)")) %>%
                           mutate(Ablation = ifelse(Ablation  == "N", "No Ablation", "Ablated")) %>%
                           mutate(Darkness = ifelse(Darkness == "N", "Light", "Dark")) %>%
                           mutate(Flow = factor(Flow), Ablation = factor(Ablation), Darkness = factor(Darkness)) %>%
                           mutate(Flow = fct_relevel(Flow, c("Still Water","Flowing Water (2 BL/s)"))) %>%
                           mutate(Ablation = fct_relevel(Ablation, c("No Ablation","Ablated"))) %>%
                           mutate(Darkness = fct_relevel(Darkness, c("Light","Dark"))) %>%
                           mutate(Flow_Ablation_Darkness = factor(paste(Flow,Ablation,Darkness,sep=", "))) %>%
                           separate(Fish, c("Fish1", "Fish2"),sep="x") %>% 
                           ### Used to remove every third point to reduce autocorrelation
                           mutate(Data_ID = paste(Year,Month,Day,Trial, sep = "_"),
                                  Date = paste(Year,Month,Day,Trial, sep = "_"),
                                  Fish_ID = paste(Year,Month,Day,Trial,Fish1, sep = "_"),
                                  Comp_ID = paste(Year,Month,Day,Trial,Fish1,Fish2, sep = "_")) %>%
                           group_by(Comp_ID) %>%
                           #filter(row_number() %% 3 == 1) %>%
                           ungroup() %>%
                           mutate(Type = "Eight")

fish_comp_data <- na.omit(fish_comp_data)

fish_comp_data <- rbind(fish_comp_data,single_comp_data)

ggplot(fish_comp_data, aes(x = Angle))+
  geom_histogram()+
  facet_wrap(~Flow_Ablation_Darkness)

sd <- 1.77

min_dist <- 0.75 #sd*0.5
max_dist <- 3 #sd*1.5

#calculate average angle to all other fish in that range
fish_comp_data_ring_angles <- fish_comp_data %>% filter(Distance >= min_dist & Distance <= max_dist) %>%
                                                 group_by(Date,Year,Month,Day,Trial,
                                                          Flow,Ablation,Darkness,Type,Flow_Ablation_Darkness,Fish1,Tailbeat_Num) %>%
                                                 summarise(Mean_Angle = rad_ang_mean(Angle),
                                                           #Should be the same number each time
                                                           Fish1_Yaw_Heading_PS = mean(Fish1_Yaw_Heading_PS, rm.na = T)) %>%
                                                 ungroup()

fish_data_nn_normal <- fish_comp_data_ring_angles %>% group_by(Flow_Ablation_Darkness) %>%
                                        mutate(Fish1_Yaw_Heading_PS = Fish1_Yaw_Heading_PS - mean(Fish1_Yaw_Heading_PS, na.rm = T)) %>%
                                        ungroup()

nn_heading_sum <- fish_data_nn_normal %>% mutate(angle_bins = round_any(Mean_Angle,pi/6)) %>%
                                                group_by(angle_bins,Flow,Ablation,Darkness,Flow_Ablation_Darkness,Type) %>%
                                                summarise(mean_turn = mean(Fish1_Yaw_Heading_PS, na.rm = T),
                                                          sd_turn = sd(Fish1_Yaw_Heading_PS, na.rm = T),
                                                          se = sd(Fish1_Yaw_Heading_PS, na.rm = T) / sqrt(n()),
                                                          n = n()) %>%
                                                ungroup() %>%
                                                na.omit() %>%
                                                group_by(Flow_Ablation_Darkness) %>%
                                                mutate(se = mean(se)) %>%
                                                ungroup()
```

Distribution of turning velocities at say 90 degrees
just make a Kolmogorov–Smirnov to compare distribution. See if distribution is different
  Not larger or smaller?
Also maybe just do a t test 
trim the edge points

too much variance, not significant

Load in Single Fish

Estimating functions

```{r}

fish_data_nn_model <- fish_data_nn_normal %>% filter(Flow == "Still Water") %>% 
                                       mutate(AblationCat = ifelse(Ablation == "Ablated",1,0),
                                              DarknessCat = ifelse(Darkness == "Dark",1,0))

heading_ps_angle_base <- lm(Fish1_Yaw_Heading_PS ~ 1, data = fish_data_nn_model)

heading_ps_angle_step <- stepAIC(heading_ps_angle_base,
                                   list(upper = ~ Mean_Angle*sin(Mean_Angle)*Ablation*Darkness, lower = ~1),
                                   direction = "both",
                                   trace = F)

Anova(heading_ps_angle_step)

manual_model <- lmer(Fish1_Yaw_Heading_PS ~ sin(Mean_Angle) + sin(Mean_Angle):Ablation + sin(Mean_Angle):Darkness + sin(Mean_Angle):Ablation:Darkness + (1|Date), 
                   data = fish_data_nn_model) 
Anova(manual_model)

#Raw data
ggplot(fish_data_nn_model,
       aes(x = Mean_Angle, y = rad2deg(Fish1_Yaw_Heading_PS), color = interaction(Ablation, Darkness, sep =", ")))+
  geom_point(alpha = 0.5)+
  facet_wrap(~ interaction(Ablation, Darkness, sep =", ")) +
  geom_line(aes(y = predict(manual_model)), size = 1)+
  theme_classic() +
  xlab("direction of neighbour (degrees)") +
  ylab("focal fish turning angle (rad/sec)") +
  scale_colour_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color,"red")) +
  scale_x_continuous(breaks = unique(nn_heading_sum$angle_bins),
                     labels = rad2deg(unique(nn_heading_sum$angle_bins))) +
  labs(color = "Sensory Conditions")

graph_model_data = nn_heading_sum %>% filter(Flow == "Still Water") %>% mutate(Mean_Angle = angle_bins)

#Means graph
ggplot(nn_heading_sum %>% filter(Flow == "Still Water"),
       aes(x = angle_bins, y = rad2deg(mean_turn), color = interaction(Ablation, Darkness, sep =", ")))+
  geom_point()+
  geom_line(aes( y = rad2deg(predict(manual_model,newdata = graph_model_data, re.form = NA)) ), size = 1)+
  geom_ribbon(aes(ymin = rad2deg(predict(manual_model,newdata = graph_model_data, re.form = NA) - se),
                  ymax = rad2deg(predict(manual_model,newdata = graph_model_data, re.form = NA) + se),
                  fill = interaction(Ablation, Darkness, sep =", ")),
              alpha = 0.25, colour = NA)+
  theme_classic() +
  ylim(-100,100) +
  xlab("direction of neighbour (degrees)") +
  ylab("focal fish turning angle speed (degrees/sec)") +
  scale_colour_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color,"red")) +
  scale_fill_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color,"red")) +
  scale_x_continuous(breaks = unique(nn_heading_sum$angle_bins),
                     labels = rad2deg(unique(nn_heading_sum$angle_bins))) +
  labs(color = "Sensory Conditions")+
  ggtitle(paste0("All Neighbors From ",min_dist," to ",max_dist))+ 
  guides(fill = FALSE)

test(emmeans(manual_model, trt.vs.ctrlk ~ Ablation*Darkness | Mean_Angle,
             at = list(Mean_Angle = c(pi/2))))

test(emmeans(manual_model, trt.vs.ctrlk ~ Ablation*Darkness | Mean_Angle,
             at = list(Mean_Angle = c(pi/2))))

plot(manual_model)
qqnorm(resid(manual_model))
#qqline(resid(groups_lmer_flow))

```

Removing underlying variation to the data?

```{r}

tank_data <- fish_comp_data %>% filter(Flow == "Still Water") %>% 
                                mutate(Round_X = round_any(Fish1_X,1),
                                       Round_Y = round_any(Fish1_Y,1)) %>%
                                group_by(Flow_Ablation_Darkness, Round_X, Round_Y) %>%
                                summarise(Mean_HPS = mean(Fish1_Yaw_Heading_PS)) %>%
                                ungroup()

ggplot(tank_data , aes(Round_X, Round_Y, fill = Mean_HPS))+
  facet_wrap(~ Flow_Ablation_Darkness)+
  geom_tile()+
  theme_classic()

```