### 2024-03-19 Fish Data Analysis/NN_Accel_Heading.qmd
Commit before merge

---
title: "NN_Accel_Heading"
format: html
editor: visual
---

## Loading in the Libraries for Everything

```{r}
library(MASS)
library(tidyverse)
library(ggplot2)
library(bbmle) #For ICtab
library(car)
library(ggpubr)
library(ggdist)
library(emmeans)
library(lme4)
library(viridis)
library(lme4)
```

```{r}
rad2deg <- function(rad) {(rad * 180) / (pi)}
deg2rad <- function(deg) {(deg * pi) / (180)}
round_any <- function(x, accuracy, f=round){f(x / accuracy) * accuracy}

deg_ang_mean <- function(x){rad2deg(atan2(mean(sin(deg2rad(x))),mean(cos(deg2rad(x)))))}
rad_ang_mean <- function(x){atan2(mean(sin(x)),mean(cos(x)))}

fold_angle_0_360_to_0_180 <- function(x){abs(abs(x-180)-180)}

fold_angle_neg_180_180_to_neg_90_90 <- function(x){ifelse(x > 90,90-abs(90-x),ifelse(x < -90, -90+abs(-90-x), x))}

min_n <- function(x,n){sort(x)[1:n]}
max_n <- function(x,n){sort(x,decreasing = TRUE)[1:n]}

nth_min <- function(x,n){sort(na.omit(x))[n]}

light_no_ab_color = "#9C9C9C"   #"#C62E65" #A1BAA4" #"#d4d7dd" 06D6A0
light_ab_color = "#f5ba9e"
dark_no_ab_color = "#5e94d4"
dark_ab_color = "#49A078"   #"#1E4371"
sim_color = "#777777"

flow_split_still = "#F59DE2"
flow_split_flow = "#5DD492"

dark_split_light = "#5e94d4"
dark_split_dark = "#011627"

tailbeat_len = 19

allowed_flow_type = c("Still Water","Flowing Water (2 BL/s)")

dot_binwidth = 0.03

text_size = 16
```

Load in the data

```{r}

single_comp_data <- read.csv("Data/Fish_Comp_Values_3D_Single.csv") %>%
                           mutate(Flow = ifelse(Flow == "0", "Still Water", "Flowing Water (2 BL/s)")) %>%
                           mutate(Ablation = ifelse(Ablation  == "S", "Single", "Jingle")) %>%
                           mutate(Darkness = ifelse(Darkness == "S", "Single", "Jingle")) %>%
                           mutate(Flow = factor(Flow), Ablation = factor(Ablation), Darkness = factor(Darkness)) %>%
                           mutate(Flow = fct_relevel(Flow, c("Still Water"))) %>%
                           mutate(Ablation = fct_relevel(Ablation, c("Single"))) %>%
                           mutate(Darkness = fct_relevel(Darkness, c("Single"))) %>%
                           mutate(Flow_Ablation_Darkness = factor(paste(Flow,Ablation,Darkness,sep=", "))) %>%
                           separate(Fish, c("Fish1", "Fish2"),sep="x") %>% 
                           # # ### Used to remove every third point to reduce autocorrelation
                           mutate(Data_ID = paste(Year,Month,Day,Trial, sep = "_"),
                                  Date = paste(Year,Month,Day,Trial, sep = "_"),
                                  Fish_ID = paste(Year,Month,Day,Trial,Fish1, sep = "_"),
                                  Comp_ID = paste(Year,Month,Day,Trial,Fish1,Fish2, sep = "_")) %>%
                           group_by(Comp_ID) %>%
                           #filter(row_number() %% 3 == 1) %>%
                           ungroup() %>%
                           mutate(Type = "Single")

fish_comp_data <- read.csv("Data/Fish_Comp_Values_3D.csv") %>%
                           mutate(Flow = ifelse(Flow == "0", "Still Water", "Flowing Water (2 BL/s)")) %>%
                           mutate(Ablation = ifelse(Ablation  == "N", "No Ablation", "Ablated")) %>%
                           mutate(Darkness = ifelse(Darkness == "N", "Light", "Dark")) %>%
                           mutate(Flow = factor(Flow), Ablation = factor(Ablation), Darkness = factor(Darkness)) %>%
                           mutate(Flow = fct_relevel(Flow, c("Still Water","Flowing Water (2 BL/s)"))) %>%
                           mutate(Ablation = fct_relevel(Ablation, c("No Ablation","Ablated"))) %>%
                           mutate(Darkness = fct_relevel(Darkness, c("Light","Dark"))) %>%
                           mutate(Flow_Ablation_Darkness = factor(paste(Flow,Ablation,Darkness,sep=", "))) %>%
                           separate(Fish, c("Fish1", "Fish2"),sep="x") %>% 
                           ### Used to remove every third point to reduce autocorrelation
                           mutate(Data_ID = paste(Year,Month,Day,Trial, sep = "_"),
                                  Date = paste(Year,Month,Day,Trial, sep = "_"),
                                  Fish_ID = paste(Year,Month,Day,Trial,Fish1, sep = "_"),
                                  Comp_ID = paste(Year,Month,Day,Trial,Fish1,Fish2, sep = "_")) %>%
                           group_by(Comp_ID) %>%
                           #filter(row_number() %% 3 == 1) %>%
                           ungroup() %>%
                           mutate(Type = "Eight")

#fish_comp_data <- rbind(fish_comp_data,single_comp_data)

x_offset <- 1.5
y_offset <- 0.75

max_x <- ceiling(max(fish_comp_data$Fish1_X)) - 1.5
min_x <- floor(min(fish_comp_data$Fish1_X)) - 1.5
max_y <- ceiling(max(fish_comp_data$Fish1_Y)) - 0.75
min_y <- floor(max(fish_comp_data$Fish1_Y)) - 0.75

#Limit away from the edge
fish_comp_data <- fish_comp_data %>% filter(Fish1_X >= left_edge & Fish1_X <= right_edge) %>%
                                     filter(Fish1_Y >= bottom_edge & Fish1_Y <= top_edge)

ggplot(fish_comp_data, aes(x = Angle))+
  geom_histogram()+
  facet_wrap(~Flow_Ablation_Darkness)

sd <- 1.77

min_dist <- 0.5 #sd*0.5
max_dist <- 2 #sd*1.5

#calculate average angle to all other fish in that range
fish_comp_data_ring_angles <- fish_comp_data %>% filter(Distance >= min_dist & Distance <= max_dist) %>%
                                                 #mutate(Mean_Angle = Angle)
                                                 group_by(Date,Year,Month,Day,Trial,
                                                          Flow,Ablation,Darkness,Type,Flow_Ablation_Darkness,Fish1,Tailbeat_Num) %>%
                                                 summarise(Mean_Angle = rad_ang_mean(Angle),
                                                           #Should be the same number each time
                                                           Fish1_Yaw_Heading_PS = mean(Fish1_Yaw_Heading_PS, rm.na = T)) %>%
                                                 ungroup()

fish_data_nn_normal <- fish_comp_data_ring_angles %>% group_by(Flow_Ablation_Darkness) %>%
                                        mutate(Fish1_Yaw_Heading_PS = Fish1_Yaw_Heading_PS - mean(Fish1_Yaw_Heading_PS, na.rm = T)) %>%
                                        ungroup()

nn_heading_sum <- fish_data_nn_normal %>% mutate(angle_bins = round_any(Mean_Angle,pi/6)) %>%
                                                group_by(angle_bins,Flow,Ablation,Darkness,Flow_Ablation_Darkness,Type) %>%
                                                summarise(mean_turn = mean(Fish1_Yaw_Heading_PS, na.rm = T),
                                                          sd_turn = sd(Fish1_Yaw_Heading_PS, na.rm = T),
                                                          se = sd(Fish1_Yaw_Heading_PS, na.rm = T) / sqrt(n()),
                                                          n = n()) %>%
                                                ungroup() %>%
                                                na.omit() %>%
                                                group_by(Flow_Ablation_Darkness) %>%
                                                mutate(se = mean(se)) %>%
                                                ungroup()
```

Distribution of turning velocities at say 90 degrees
just make a Kolmogorov–Smirnov to compare distribution. See if distribution is different
  Not larger or smaller?
Also maybe just do a t test 
trim the edge points

too much variance, not significant

Load in Single Fish

Estimating functions

Kolmogorov–Smirnov around 90 degrees

```{r}

perp_angle_df <- fish_data_nn_normal %>% filter(Mean_Angle >= deg2rad(80) & Mean_Angle <= deg2rad(100)) %>% filter(Flow == "Still Water")#%>% filter(Distance >= 0 & Distance <= 3) %>% filter(Flow == "Still Water")

perp_angle_df_neg <- fish_data_nn_normal %>% filter(Mean_Angle >= deg2rad(-100) & Mean_Angle <= deg2rad(-80)) %>% filter(Flow == "Still Water") %>% mutate(Fish1_Yaw_Heading_PS = -1*Fish1_Yaw_Heading_PS)

perp_angle_df <- rbind(perp_angle_df,perp_angle_df_neg)

perp_angle_df <- perp_angle_df %>% group_by(Flow,Ablation,Darkness,Flow_Ablation_Darkness,Type) %>%
                                                mutate(mean_turn = mean(Fish1_Yaw_Heading_PS, na.rm = T),
                                                       sd_turn = sd(Fish1_Yaw_Heading_PS, na.rm = T),
                                                       n = n()) %>%
                                                ungroup()

#single_fish_df <- perp_angle_df %>% filter(Darkness == "Single" & Ablation == "Single")
LNDN_fish_df <- perp_angle_df %>% filter(Darkness == "Light" & Ablation == "No Ablation")
LYDN_fish_df <- perp_angle_df %>% filter(Darkness == "Light" & Ablation == "Ablated")
LNDY_fish_df <- perp_angle_df %>% filter(Darkness == "Dark" & Ablation == "No Ablation")
LYDY_fish_df <- perp_angle_df %>% filter(Darkness == "Dark" & Ablation == "Ablated")

# LNDN_KS <- ks.test(single_fish_df$Fish1_Yaw_Heading_PS,LNDN_fish_df$Fish1_Yaw_Heading_PS)
# LYDN_KS <- ks.test(single_fish_df$Fish1_Yaw_Heading_PS,LYDN_fish_df$Fish1_Yaw_Heading_PS)
# LNDY_KS <- ks.test(single_fish_df$Fish1_Yaw_Heading_PS,LNDY_fish_df$Fish1_Yaw_Heading_PS)
# LNDY_KS <- ks.test(single_fish_df$Fish1_Yaw_Heading_PS,LYDY_fish_df$Fish1_Yaw_Heading_PS)

#p.adjust(c(LNDN_KS$p.value,LYDN_KS$p.value,LNDY_KS$p.value,LNDY_KS$p.value))

perp_model <- aov(Fish1_Yaw_Heading_PS ~ Darkness*Ablation, data = perp_angle_df)
Anova(perp_model)
qqnorm(resid(perp_model))
TukeyHSD(perp_model)

ggplot(perp_angle_df, aes(x = interaction(Ablation, Darkness, sep =", "), y = rad2deg(Fish1_Yaw_Heading_PS),
                          fill = interaction(Ablation, Darkness, sep =", "), color = interaction(Ablation, Darkness, sep =", ")))+
  stat_dots(side = "both", binwidth = unit(dot_binwidth, "npc")) +
  geom_point(aes(y = rad2deg(mean_turn)), size = 5, color = "black") +
  geom_errorbar(aes(ymin=rad2deg(mean_turn-sd_turn), ymax=rad2deg(mean_turn+sd_turn)), width=.2, color = "black")+
  theme_classic()+
  ggtitle(paste0("All Neighbors From ",min_dist," to ",max_dist))+
  scale_fill_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color)) +
  scale_color_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color)) +
  theme(legend.position = "none")+
  xlab("")+
  ylab("Turning Angle (Degrees / s)")

```

```{r}

fish_data_nn_model <- fish_data_nn_normal %>% filter(Flow == "Still Water") %>% 
                                       mutate(AblationCat = ifelse(Ablation == "Ablated",1,0),
                                              DarknessCat = ifelse(Darkness == "Dark",1,0)) %>% na.omit()

heading_ps_angle_base <- lm(Fish1_Yaw_Heading_PS ~ 1, data = fish_data_nn_model)

heading_ps_angle_step <- stepAIC(heading_ps_angle_base,
                                   list(upper = ~ Mean_Angle*sin(Mean_Angle)*Ablation*Darkness, lower = ~1),
                                   direction = "both",
                                   trace = F)

Anova(heading_ps_angle_step)

manual_model <- lmer(Fish1_Yaw_Heading_PS ~ sin(Mean_Angle) + sin(Mean_Angle):Ablation + sin(Mean_Angle):Darkness + sin(Mean_Angle):Ablation:Darkness + (1|Date), 
                   data = fish_data_nn_model) 
Anova(manual_model)

#Raw data
ggplot(fish_data_nn_model,
       aes(x = Mean_Angle, y = rad2deg(Fish1_Yaw_Heading_PS), color = interaction(Ablation, Darkness, sep =", ")))+
  geom_point(alpha = 0.5)+
  facet_wrap(~ interaction(Ablation, Darkness, sep =", ")) +
  geom_line(aes(y = predict(manual_model)), size = 1)+
  theme_classic() +
  xlab("direction of neighbour (degrees)") +
  ylab("focal fish turning angle (rad/sec)") +
  scale_colour_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color,"red")) +
  scale_x_continuous(breaks = unique(nn_heading_sum$angle_bins),
                     labels = rad2deg(unique(nn_heading_sum$angle_bins))) +
  labs(color = "Sensory Conditions")

graph_model_data = nn_heading_sum %>% filter(Flow == "Still Water") %>% mutate(Mean_Angle = angle_bins)

#Means graph
ggplot(nn_heading_sum %>% filter(Flow == "Still Water"),
       aes(x = angle_bins, y = rad2deg(mean_turn), color = interaction(Ablation, Darkness, sep =", ")))+
  geom_point()+
  geom_line(aes( y = rad2deg(predict(manual_model,newdata = graph_model_data, re.form = NA)) ), size = 1)+
  # geom_ribbon(aes(ymin = rad2deg(predict(manual_model,newdata = graph_model_data, re.form = NA) - se),
  #                 ymax = rad2deg(predict(manual_model,newdata = graph_model_data, re.form = NA) + se),
  #                 fill = interaction(Ablation, Darkness, sep =", ")),
  #             alpha = 0.25, colour = NA)+
  theme_classic() +
  ylim(-100,100) +
  xlab("direction of neighbour (degrees)") +
  ylab("focal fish turning angle speed (degrees/sec)") +
  scale_colour_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color,"red")) +
  scale_fill_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color,"red")) +
  scale_x_continuous(breaks = unique(nn_heading_sum$angle_bins),
                     labels = rad2deg(unique(nn_heading_sum$angle_bins))) +
  labs(color = "Sensory Conditions")+
  ggtitle(paste0("All Neighbors From ",min_dist," to ",max_dist))+ 
  guides(fill = FALSE)

test(emmeans(manual_model, trt.vs.ctrlk ~ Ablation*Darkness | Mean_Angle,
             at = list(Mean_Angle = c(pi/2))))

test(emmeans(manual_model, trt.vs.ctrlk ~ Ablation*Darkness | Mean_Angle,
             at = list(Mean_Angle = c(pi/2))))

plot(manual_model)
qqnorm(resid(manual_model))
#qqline(resid(groups_lmer_flow))

```

Removing underlying variation to the data?

```{r}

tank_data <- fish_comp_data %>% filter(Flow == "Still Water") %>% 
                                mutate(Round_X = round_any(Fish1_X,1),
                                       Round_Y = round_any(Fish1_Y,1)) %>%
                                group_by(Flow_Ablation_Darkness, Round_X, Round_Y) %>%
                                summarise(Mean_HPS = mean(Fish1_Yaw_Heading_PS)) %>%
                                ungroup()

ggplot(tank_data , aes(Round_X, Round_Y, fill = Mean_HPS))+
  facet_wrap(~ Flow_Ablation_Darkness)+
  geom_tile()+
  theme_classic()

```

When are single fish flat?
```{r}

x_offset <- 1.5
y_offset <- 0.75

top_edge <- 3 - y_offset 
bottom_edge <- 0 + y_offset
left_edge <- 0 + x_offset
right_edge <- 7 - x_offset

#calculate average angle to all other fish in that range
single_fish_mean_angles <- single_comp_data %>%   #filter(Flow_Ablation_Darkness == "Still Water, No Ablation, Light") %>%
                                                filter(Fish1_X >= left_edge & Fish1_X <= right_edge) %>%
                                                filter(Fish1_Y >= bottom_edge & Fish1_Y <= top_edge) %>%
                                                #filter(Distance >= min_dist & Distance <= max_dist) %>%
                                                group_by(Date,Year,Month,Day,Trial,
                                                          Flow,Ablation,Darkness,Type,Flow_Ablation_Darkness,Fish1,Tailbeat_Num) %>%
                                                 # summarise(Mean_Angle = rad_ang_mean(Angle),
                                                 #           #Should be the same number each time
                                                 #           Fish1_Yaw_Heading_PS = mean(Fish1_Yaw_Heading_PS, rm.na = T),
                                                 #           Fish1_X = mean(Fish1_X, rm.na = T),
                                                 #           Fish1_Y = mean(Fish1_Y, rm.na = T)) %>%
                                                 ungroup() %>%
                                                 #normalize
                                                 mutate(Fish1_Yaw_Heading_PS = Fish1_Yaw_Heading_PS - mean(Fish1_Yaw_Heading_PS, na.rm = T),
                                                        Mean_Angle = Angle)

single_flat_lm <- lm(Fish1_Yaw_Heading_PS ~ 1, data = single_fish_mean_angles)
single_lin_lm <- lm(Fish1_Yaw_Heading_PS ~ Mean_Angle, data = single_fish_mean_angles)
single_sin_lm <- lm(Fish1_Yaw_Heading_PS ~ sin(Mean_Angle), data = single_fish_mean_angles)

AIC(single_flat_lm,single_lin_lm,single_sin_lm) %>% arrange(AIC)

single_perp_angle_df <- single_fish_mean_angles %>% filter(Mean_Angle >= deg2rad(80) & Mean_Angle <= deg2rad(100)) %>% filter(Flow == "Still Water")

single_perp_angle_df_neg <- single_fish_mean_angles %>% filter(Mean_Angle >= deg2rad(-100) & Mean_Angle <= deg2rad(-80)) %>% filter(Flow == "Still Water") %>% mutate(Fish1_Yaw_Heading_PS = -1*Fish1_Yaw_Heading_PS)

single_perp_angle_df <- rbind(single_perp_angle_df,single_perp_angle_df_neg)

t.test(single_perp_angle_df$Fish1_Yaw_Heading_PS)

effect_size <- mean(single_perp_angle_df$Fish1_Yaw_Heading_PS) / sd(single_perp_angle_df$Fish1_Yaw_Heading_PS)
effect_size

ggplot(single_perp_angle_df, aes(x = Fish1_Yaw_Heading_PS))+
  geom_histogram(color = "black", fill = "red", binwidth = 1)+
  theme_classic()

ggplot(single_fish_mean_angles, aes(x = Mean_Angle, y = Fish1_Yaw_Heading_PS))+
  geom_point(color = "red")+
  # geom_smooth(method = "glm", formula = y ~ 1, color = "black")+
  # geom_smooth(method = "glm", formula = y ~ x, color = "green")+
  geom_smooth(method = "glm", formula = y ~ sin(x), color = "blue")+
  theme_classic()

single_fish_mean_angles_heatmap <- single_fish_mean_angles %>% mutate(RoundX = round_any(Fish1_X, 0.5),
                                                                      RoundY = round_any(Fish1_Y, 0.5)) %>%
                                                               group_by(RoundX,RoundY) %>%
                                                               summarise(mean_turn = mean(Mean_Angle),
                                                                         n = n()) %>%
                                                               ungroup()

ggplot(single_fish_mean_angles_heatmap, aes(x = RoundX, y = RoundY, fill = mean_turn))+
  geom_tile()+
  coord_equal()+
  scale_fill_gradient2(low = "blue", mid = "white", high = "red")+
  theme_classic()

possible_offsets <- expand_grid(x = seq(0,2,0.25), y = seq(0,1.25,0.25))

offset_return <- function(x,y){
  te <- 3 - y 
  be <- 0 + y
  le <- 0 + x
  re <- 7 - x

  tester_data <- single_comp_data %>%  filter(Fish1_X >= le & Fish1_X <= re) %>%
                                        filter(Fish1_Y >= be & Fish1_Y <= te) %>%
                                        mutate(Fish1_Yaw_Heading_PS = Fish1_Yaw_Heading_PS - mean(Fish1_Yaw_Heading_PS, na.rm = T))

  tester_perp_df <- tester_data %>% filter(Angle >= deg2rad(80) & Angle <= deg2rad(100)) %>%
                                                      filter(Flow == "Still Water")

  tester_perp_df_neg <- tester_data %>% filter(Angle >= deg2rad(-100) & Angle <= deg2rad(-80)) %>% 
                                                        filter(Flow == "Still Water") %>% 
                                                        mutate(Fish1_Yaw_Heading_PS = -1*Fish1_Yaw_Heading_PS)

  tester_perp_df <- rbind(tester_perp_df,tester_perp_df_neg)

  t_test_out <- t.test(tester_perp_df$Fish1_Yaw_Heading_PS)

  return(t_test_out$p.val)
}

for(x in seq(0,2,0.25)){
  for (y in seq(0,1.25,0.25)) {
    print(paste(x,y))
    print(offset_return(x,y))
  }
}

```